### 2025-01-20 paper.tex
Write paper.

\documentclass{article}

\usepackage[style=alphabetic]{biblatex}
\addbibresource{paper.bib}

% https://ncatlab.org/nlab/show/Yoneda+embedding#ReferencesNotation
\usepackage[utf8]{inputenc}
\DeclareFontFamily{U}{min}{}
\DeclareFontShape{U}{min}{m}{n}{<-> udmj30}{}
\newcommand\yo{\!\text{\usefont{U}{min}{m}{n}\symbol{'207}}\!}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{mathpartir}
\usepackage{tcolorbox}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{example}[theorem]{Example}

\newcommand*{\Prop}{\mathbf{Prop}}
\newcommand*{\Set}{\mathbf{Set}}
\newcommand*{\Ring}{\mathbf{Ring}}
\newcommand*{\CRing}{\mathbf{CRing}}
\newcommand*{\Graph}{\mathbf{Graph}}
\newcommand*{\Sh}{\mathrm{Sh}}
\newcommand*{\Psh}{\mathrm{Psh}}
\newcommand*{\B}{\mathbb{B}}
\newcommand*{\C}{\mathcal{C}}
\newcommand*{\N}{\mathbb{N}}
\newcommand*{\T}{\mathbb{T}}
\newcommand*{\Z}{\mathbb{Z}}
\newcommand*{\apart}{\mathrel{\#}}

% https://tex.stackexchange.com/a/223246
\newcommand*{\wand}{\mathbin{-\mkern-6mu*}}

\title{Instantiating Bunched Dependent Type Theory for Monoidal Classifying Toposes}
\author{finegeometer}

\begin{document}
\maketitle

\begin{abstract}
    I provide a simple condition on a geometric theory \(\T\),
    which suffices to interpret the bunched type theory \(\mathbf{BT}(*, 1, \Sigma, \Pi, \Pi^*)\) of \cite{schopp} into the classifying topos \(\Set[\T]\).
\end{abstract}

\section{Intro}

\cite{schopp} introduces a bunched dependent type theory, \(\mathbf{BT}(*, 1, \Sigma, \Pi, \Pi^*)\),
of potentially general utility. But they only instantiate it for a single topos: the Schanuel topos.

By stitching together a few results from the literature,
I provide an easy method to verify that \(\mathbf{BT}(*, 1, \Sigma, \Pi, \Pi^*)\)
can be interpreted in a classifying topos of your choice,
so you can start \emph{using} the language as quickly as possible.

\section{Main Theorem}

\begin{theorem}
    Consider a geometric theory, \(\T\).
    For each pair \(A,B\) of sorts of \(\T\), let \((\sim) : A \times B \to \Prop\) be a relation, defined in terms of \(\T\)'s signature.
    Further suppose that \((\sim)\) is symmetric, and that the following holds.

    \begin{quote}
        Let \(X\) be an arbitrary sort of \(\T\), and let \(x : X\).
        Then if you restrict each sort \(A\) of \(\T\) to the subset \(\{a : A \mid a \sim x\}\),
        the result is another \(\T\)-model.
    \end{quote}

    \noindent Then \(\mathbf{BT}(*, 1, \Sigma, \Pi, \Pi^*)\) interprets into \(\Set[\T]\).
\end{theorem}

\subsection{Example}

As an example, let \(\T\) be the theory of local rings, and let \(\sim : R \times R \to \Prop\) be the relation \(a \sim b \iff a \cdot b = b \cdot a\).

\((\sim)\) is clearly symmetric. And for any \(x : R\), \(\{a : R \mid a \sim x\}\) is a local subring of \(R\).
So the conditions of the theorem are satisfied, and therefore the type theory interprets into \(\Set[\T]\).

\subsection{Details} \label{Details}

Some of the concepts in the theorem statement may be insufficiently precise. I address that here.

To say \(\sim\) is symmetric is to say that for every pair \(A,B\) of sorts in \(\T\),
the following is provable in the language of the theory.
\[a : A, b : B \vdash a \sim b \implies b \sim a\]

To say that the subsets \(\{a : A \mid a \sim x\}\) form another \(\T\)-model is to say the following:
\begin{itemize}
    \item For every sort \(X\) and every function symbol \(f : A_1 \times \dots \times A_n \to B\), the following is provable:
    \[x : X, a_1 : A_1 \dots a_n : A_n, a_1 \sim x \dots a_n \sim x \vdash f(a_1 \dots a_n) \sim x\]
    \item For every sort \(X\) and every axiom \(a_1 : A_1 \dots a_n : A_n, P \vdash Q\):

    Let \(P'\) be produced from \(P\)
    by replacing each existential quantifier \(\exists b : B, \dots\)
    with \(\exists b : B, b \sim x \land \dots\).
    Similarly define \(Q'\) in terms of \(Q\).
    Then the following is provable:
    \[x : X, a_1 : A_1 \dots a_n : A_n, a_1 \sim x \dots a_n \sim x, P' \vdash Q'\]
\end{itemize}

To see how this plays out in practice, consider the example from above --- the theory of local rings.

\begin{tcolorbox}[colback=red!5!white, colframe=red!50!black]
    \begin{mathpar}
        R \text{ sort}
        \and 0 : R
        \and 1 : R
        \\ (+) : R \times R \to R
        \and (\cdot) : R \times R \to R
        \and (-) : R \to R
        \\ a,b,c : R \vdash (a + b) + c = a + (b + c)
        \and a,b : R \vdash a + b = b + a
        \\ a : R \vdash a + 0 = a
        \and a : R \vdash a + (-a) = 0
        \\ a,b,c : R \vdash (a \cdot b) \cdot c = a \cdot (b \cdot c)
        \and a : R \vdash a \cdot 1 = a
        \\ a,b,c : R \vdash (a + b) \cdot c = a \cdot c + b \cdot c
        \and a,b,c : R \vdash a \cdot (b + c) = a \cdot b + a \cdot c
        \\ 0 = 1 \vdash \bot
        \and a,b : R, a + b = 1 \vdash
            (\exists c, a \cdot c = 1 \land c \cdot a = 1) \lor
            (\exists c, b \cdot c = 1 \land c \cdot b = 1)
    \end{mathpar}
    \tcblower
    \[a \sim b \iff a \cdot b = b \cdot a\]
\end{tcolorbox}

The sub-\(\T\)-model conditions say the following:
\begin{tcolorbox}[colback=blue!5!white, colframe=blue!50!black]
    \begin{align*}
        x : R &\vdash 0 x = x 0
        \\ x : R &\vdash 1 x = x 1
        \\ x,a,b : R, a x = x a, b x = x b &\vdash (a + b) x = x (a + b)
        \\ x,a,b : R, a x = x a, b x = x b &\vdash (a b) x = x (a b)
        \\ x,a : R, a x = x a &\vdash (-a) x = x (-a)
    \end{align*}
    \tcblower
    \begin{align*}
        x,a,b,c : R, ax=xa, bx=xb, cx=xc &\vdash (a + b) + c = a + (b + c)
        \\ x,a,b : R, ax=xa, bx=xb &\vdash a + b = b + a
        \\ x,a : R, ax=xa &\vdash a + 0 = a
        \\ x,a : R, ax=xa &\vdash a + (-a) = 0
        \\ x,a,b,c, ax=xa, bx=xb, cx=xc : R &\vdash (ab)c = a(bc)
        \\ x,a : R, ax=xa &\vdash a1 = a
        \\ x,a,b,c, ax=xa, bx=xb, cx=xc : R &\vdash (a + b) c = a c + b c
        \\ x,a,b,c, ax=xa, bx=xb, cx=xc : R &\vdash a (b + c) = a b + a c
        \\ x : R, 0 = 1 &\vdash \bot
        \\ x,a,b : R, ax=xa, bx=xb, a + b = 1 &\vdash \\
            (\exists c, cx=xc \land ac = 1 \land ca = 1) &\lor
            (\exists c, cx=xc \land bc = 1 \land cb = 1)        
    \end{align*}
\end{tcolorbox}

All of these are easily verified. More to the point, they capture the idea
that for all \(x : R\), the subset \(\{a : R \mid ax = xa\}\) is a local ring.

\section{Proof of Main Theorem}

To prove the main theorem, begin by recalling some results from \cite{schopp}.

\begin{lemma}
    The type theory \(\mathbf{BT}(*, 1, \Sigma, \Pi, \Pi^*)\) can be modeled in any \\
    \((*, 1, \Sigma, \Pi, \Pi^*)\)-type-category.
\end{lemma}

\begin{proof}
    \cite{schopp}, chapter 6.
\end{proof}

\begin{lemma}
    Let \(\B\) be a topos equipped with a strict affine symmetric monoidal closed structure \((*,\wand)\),
    such that \(- * A\) preserves pullbacks for all \(A\).
    Further suppose we have a canonical choice of pullback for every span in \(\B\).
    Then the family fibration over \(\B\) is a \((*, 1, \Sigma, \Pi, \Pi^*)\)-type-category.
\end{lemma}

\begin{proof}
    \cite{schopp}, section 3.5. Since we are in a topos, the condition that the monomorphism \(A * B \hookrightarrow A \times B\) is strong is trivial.
\end{proof}

Next, specialize to sheaf toposes.

\begin{lemma} \label{sh}
    Let \((C,J)\) be a (small) site, such that \(C\) is finitely complete,
    with a canonical choice of pullbacks.
    Let \(*\) be a strict affine symmetric monoidal structure on \(C\) that preserves pullbacks and covers.
    Then \(\Sh(C,J)\) satisfies the conditions of the above lemma.
\end{lemma}

\begin{proof}
    \(\Sh(C,J)\) is a Grothendieck topos, hence a topos. There is a canonical choice of pullbacks in \(\Sh(C,J)\), given by pointwise lifting the canonical pullbacks in \(C\).

    Now, theorem 4.3.2 from \cite{biering} shows the existence of a monidal structure \(\otimes^\Sh\) on \(\Sh(C,J)\),
    given by transporting Day convolution across the adjunction \((\mathbf{a} \dashv i) : \Psh(C,J) \rightleftarrows \Sh(C,J)\).
    Corollary 4.3.10 from the same shows that it is closed.
    Chasing definitions quickly shows \(\otimes^\Sh\) is symmetric.
    And the monoidal unit is the sheafification of the Yoneda embedding of \(*\)'s unit;
    since \(*\) is affine, this simplifies to the terminal object.

    In other words, \(\otimes^\Sh\) is an affine symmetric monoidal closed structure.
    It remains to show that this is strict --- that the canonical map \(A \otimes^\Sh B \to A \times B\) in \(\Sh(C,J)\) is always a monomorphism.

    We can reduce to the case of presheaves.
    Letting \(\otimes^\Psh\) represent Day convolution on \(\Psh(C)\),
    the map of interest is just the result of applying the sheafification functor, which preserves monomorphisms, to the canonical map \(iA \otimes^\Psh iB \to iA \times iB\).

    Expanding both \(\otimes^\Psh\) and \(\times\) as Day convolutions,
    this is a set-map, natural in \(c \in C^\mathrm{op}\), of the following type.
    \begin{scriptsize}\[\left(\int^{c_1,c_2 \in C} A(c_1) \times B(c_2) \times (c \xrightarrow{C} c_1 * c_2)\right) \to \left(\int^{c_1,c_2 \in C} A(c_1) \times B(c_2) \times (c \xrightarrow{C} c_1 \times c_2)\right)\]\end{scriptsize}
    This map is exactly what you'd expect it to be; it composes the \(c \xrightarrow{C} c_1 * c_2\) component
    with the canonical map \(c_1 * c_2 \xrightarrow{C} c_1 \times c_2\), and leaves everything else alone.
    Since that canonical map is mono, this is too.
\end{proof}

Finally, set the site to a syntactic site, to understand classifying toposes.

\begin{lemma}
    Let \(\T\) be a geometric theory, satisfying the conditions of the main theorem.
    Then the syntactic site for \(\T\) satisfies the conditions of the above lemma.
\end{lemma}

\begin{proof}
    Let us define a few shorthands.

    We'll allow ourselves to write sequents involving compound types, such as \(p : A \times B \vdash \pi_1 p \sim \pi_2 p\).
    These can be straightforwardly ``compiled out'' to ordinary sequents, such as \(a : A, b : B \vdash a \sim b\).

    Next, if we have a context \(\Gamma = (a_1 : A_1, \dots, a_n : A_n, \phi_1 \dots \phi_p)\),
    we'll reuse the name \(\Gamma\) for the type \(\{(a_1, \dots, a_n) : A_1 \times \dots \times A_n \mid \phi_1 \land \dots \land \phi_p\}\).
    If we have a substitution \(f : \Gamma \to \Delta\) and a term \(\gamma : \Gamma\),
    there is a natural way to define the term \(f\gamma : \Delta\).

    Finally, if \(\gamma = (a_1, \dots, a_m) : \Gamma\) and \(\delta = (b_1, \dots, b_n) : \Delta\),
    we write \(\gamma \sim \delta\) as a shorthand for \(\bigwedge_{i=1}^m \bigwedge_{j=1}^n a_i \sim b_j\).

    With that out of the way, let's begin the proof.

    \pagebreak

    The syntactic category is guaranteed to be finitely complete, with a canonical construction for pullbacks.
    Specifically, the empty context is terminal,
    and the pullback of the span \(\Gamma \xrightarrow{f} \Xi \xleftarrow{g} \Delta\) is the context \((\gamma : \Gamma, \delta : \Delta, f\gamma = g\delta)\).

    Given contexts \(\Gamma\) and \(\Delta\), define \(\Gamma * \Delta = (\gamma : \Gamma, \delta : \Delta, \gamma \sim \delta)\).
    I claim this is a strict affine symmetric monoidal product on the syntactic category.
    \begin{itemize}
        \item Associativity and the unit laws are easily checked, and symmetry reduces to the symmetry of \(\sim\).
        \item Functoriality says that if \(f : \Gamma \to \Gamma'\), \(f \times \mathrm{id}_\Delta\) is a valid map \(\Gamma * \Delta \to \Gamma' * \Delta\).
        This reduces to the claim \(x : \Delta, \gamma : \Gamma, \gamma \sim x \vdash f\gamma \sim x\).
        Further reducing to the same claim about the basic sorts and function symbols, this is exactly one of the hypotheses, as detailed in Section \ref{Details}.
        \item Since \(\Gamma * \Delta\) is a subtype of \(\Gamma \times \Delta\), strictness holds, and the coherence equations are trivial.
    \end{itemize}

    It remains to show that for any context \(\Theta\), the functor \(- * \Theta\) preserves pullbacks and covers.

    Preservation of pullbacks is a direct computation. For the span \(\Gamma \xrightarrow{f} \Xi \xleftarrow{g} \Delta\),
    it simply says the following two contexts are isomorphic, a fact which is easily checked.
    \begin{align*}
        &(\gamma : \Gamma, \delta : \Delta, f\gamma = g\delta, \theta : \Theta, \gamma \sim \theta, \delta \sim \theta) \\
        &(\gamma : \Gamma, \theta_1 : \Theta, \gamma \sim \theta_1, \delta : \Delta, \theta_2 : \Theta, \delta \sim \theta_2, f\gamma = g\delta, \theta_1 = \theta_2)
    \end{align*}

    That leaves covers. Suppose \(\{f_i : \Gamma_i \to \Delta \mid i \in I\}\) is a cover,
    meaning \(\delta : \Delta \vdash \bigvee_{i \in I} \exists \gamma : \Gamma_i, f_i\gamma = \delta\) is \(\T\)-provable.

    We know by assumption that for any sort \(X\) and any \(x : X\),
    \(\T\)-provable statements remain \(\T\)-provable upon restricting each sort \(A\) to its subset \(\{a : A \mid a \sim x\}\).
    Iterating this fact, the same is true if we replace the sort \(X\) by a context, such as \(\Theta\).

    Therefore, we know the following statement is provable.
    \[\theta : \Theta, \delta : \{\delta : \Delta \mid \delta \sim \theta\} \vdash \bigvee_{i \in I} \exists \gamma : \{\gamma : \Gamma_i \mid \gamma \sim \theta\}, f_i\gamma = \delta\]

    But that's equivalent to the following.
    \[\delta : (\Delta * \Theta) \vdash \bigvee_{i \in I} \exists \gamma : (\Gamma_i * \Theta), (f_i * \mathrm{id}_\Theta)\gamma = \delta\]

    So \(\{f_i * \mathrm{id}_\Theta : \Gamma_i * \Theta \to \Delta * \Theta \mid i \in I\}\) is a cover, as required.
\end{proof}

Putting everything together:

\begin{proof}[Proof of Main Theorem]
    Combine the above four lemmas.
\end{proof}

\section{Using the Monoidal Structure}

For practical use, the main theorem isn't quite enough.
You'll need to know not just that \(*\)-types exist, but how they work.
If \(R\) is the generic local ring, you'll want to know that \(R * R = \{(a,b) : R^2 \mid ab = ba\}\).

Hence this next theorem.

\begin{theorem}
    Recall that \(\Set[\T]\) contains the generic model of \(\T\).
    Suppose \(A\) and \(B\) are sorts of \(\T\), interpreted in the generic model.
    Then, internally to the language, a pair \((a,b)\) is in the image of the inclusion \(A * B \hookrightarrow A \times B\) exactly when \(a \sim b\).
\end{theorem}

\begin{proof}
    To avoid confusion, write \(A\) to refer to the sort in \(\T\),
    \((a:A)\) to refer to the \emph{context} in \(\T\),
    and \(U_A\) to refer to the corresponding type in the generic model.

    Then \(U_A\) interprets as \(\mathbf{a}\yo(a:A)\), and similarly for \(B\).
    The type \(\{(a,b) : U_A \times U_B \mid a \sim b\}\) interprets as \(\mathbf{a}\yo(a:A, b:B, a \sim b)\),
    and its inclusion into \(U_A \times U_B\) is \(\mathbf{a} \circ \yo\) applied to the obvious substitution \((a:A,b:B,a \sim b) \to (a:A,b:B)\).

    Now, in the proof of Lemma \ref{sh}, we computed the interpretation of a general inclusion \(X * Y \hookrightarrow X \times Y\).
    Specializing to the types \(U_A\) and \(U_B\), that result simplifies to the obvious map
    \(\mathbf{a}\yo((a:A) * (b:B)) \to \mathbf{a}\yo((a:A) \times (b:B))\).
    That is, it's \(\mathbf{a} \circ \yo\) applied to the obvious substitution from \((a:A,b:B,a \sim b)\) to \((a:A,b:B)\).

    So the interpretations match. Thus \(U_A * U_B = \{(a,b) : U_A \times U_B \mid a \sim b\}\), as subsets of \(U_A \times U_B\).
\end{proof}

\section{Future Work}

The purpose of this paper is to \emph{enable} future work.
Namely, the exploration of monoidal toposes via their internal languages.

But beyond that, there are a few other directions to pursue.

First, \cite{blechschmidt} provides a powerful reasoning principle for the generic model of an arbitrary theory.
It would be useful to either generalize or leverage this to prove facts involving the monoidal structure.
For instance, calculations in the interpretation sugggest that the generic local ring should support
the isomorphism \((\{x : R | x^2 = 0\} \wand R) \simeq R^2\),
analogous to the Kock-Lawvere axiom \((\{x : R | x^2 = 0\} \to R) \simeq R^2\)
supported by the generic \emph{commutative} local ring.
The commutative case follows immediately from \cite[Theorem 4.10]{blechschmidt};
is there a similarly slick way to derive the noncommutative version?

Second, it would be worthwhile to extend this result to other type theories.
For instance, \cite{schopp} defines the type theory
\(\mathbf{BT}(*, 1, \Sigma, \Pi, \Pi^*, B^{*(M:A)}, \Sigma^*)\),
which provides monoidal sigma types as well as pi types.
Furthermore, it appears that \cite{schopp}'s type theories are not the end of the story,
based both on their comments in section 13 and the existence of many competing approaches in the literature.
As new and better type theories are designed, it will be valuable to have a quick way to know when they apply.

\printbibliography

\end{document}

### 2025-01-21 paper.tex
Assume less familiarity with Schopp's work.

@@
 \begin{abstract}
-    I provide a simple condition on a geometric theory \(\T\),
-    which suffices to interpret the bunched type theory \(\mathbf{BT}(*, 1, \Sigma, \Pi, \Pi^*)\) of \cite{schopp} into the classifying topos \(\Set[\T]\).
+    In their PhD thesis, Sch\"opp introduces a type theory designed to capture the monoidal structure in the Schanuel topos.
+    I provide a simple condition that can be used to verify whether the type theory interprets into a classifying topos of your choice.
 \end{abstract}
@@
 
-To prove the main theorem, begin by recalling some results from \cite{schopp}.
-
-\begin{lemma}
-    The type theory \(\mathbf{BT}(*, 1, \Sigma, \Pi, \Pi^*)\) can be modeled in any \\
-    \((*, 1, \Sigma, \Pi, \Pi^*)\)-type-category.
-\end{lemma}
-
-\begin{proof}
-    \cite{schopp}, chapter 6.
-\end{proof}
+To prove the main theorem, we begin by recalling a result from \cite{schopp}.
 
@@
     Further suppose we have a canonical choice of pullback for every span in \(\B\).
-    Then the family fibration over \(\B\) is a \((*, 1, \Sigma, \Pi, \Pi^*)\)-type-category.
+    Then the type theory \(\mathbf{BT}(*, 1, \Sigma, \Pi, \Pi^*)\) interprets into \(\B\).
 \end{lemma}
@@
 \begin{proof}
-    \cite{schopp}, section 3.5. Since we are in a topos, the condition that the monomorphism \(A * B \hookrightarrow A \times B\) is strong is trivial.
+    Chapter 6 of \cite{schopp} shows that we can interpret the type theory in any ``\((*, 1, \Sigma, \Pi, \Pi^*)\)-type-category''.
+    Section 3.5 tells us that \(\B\) is such a category.
 \end{proof}
 
-Next, specialize to sheaf toposes.
+Next, we specialize to sheaf toposes.
 
@@
 
-Finally, set the site to a syntactic site, to understand classifying toposes.
+Finally, we set the site to a syntactic site, to understand classifying toposes.
 
@@
 
-    \pagebreak
+    \vspace{1em}
 
-    The syntactic category is guaranteed to be finitely complete, with a canonical construction for pullbacks.
+    \noindent The syntactic category is guaranteed to be finitely complete, with a canonical construction for pullbacks.
     Specifically, the empty context is terminal,

### 2025-06-11 old/1.tex
minor rephrase

% % References:
% % schopp: https://ulrichschoepp.de/Docs/th.pdf
% % biering: https://ncatlab.org/nlab/files/Biering-BunchedLogic.pdf
% % blechschmidt: https://rawgit.com/iblech/internal-methods/master/paper-qcoh.pdf

\documentclass{article}

% https://ncatlab.org/nlab/show/Yoneda+embedding#ReferencesNotation
\usepackage[utf8]{inputenc}
\DeclareFontFamily{U}{min}{}
\DeclareFontShape{U}{min}{m}{n}{<-> udmj30}{}
\newcommand\yo{\!\text{\usefont{U}{min}{m}{n}\symbol{'207}}\!}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{mathpartir}
\usepackage{quiver}

\usepackage{todonotes}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{example}[theorem]{Example}

\newcommand*{\Set}{\mathbf{Set}}
\newcommand*{\Ring}{\mathbf{Ring}}
\newcommand*{\CRing}{\mathbf{CRing}}
\newcommand*{\Graph}{\mathbf{Graph}}
\newcommand*{\Sh}{\mathrm{Sh}}
\newcommand*{\Psh}{\mathrm{Psh}}
\newcommand*{\T}{\mathbb{T}}
\newcommand*{\B}{\mathbb{B}}
\newcommand*{\Z}{\mathbb{Z}}
\newcommand*{\C}{\mathcal{C}}

% https://tex.stackexchange.com/a/223246
\newcommand*{\wand}{\mathbin{-\mkern-6mu*}}

\title{Instantiating Bunched Type Theory for Monoidal Classifying Toposes}
\author{}

\begin{document}
\maketitle

\begin{abstract}
    I provide a simple condition for the classifying topos of a theory
    to be interpretable in the bunched type theory of \cite{schopp}.
    This provides a natural way for the internal logic to work with a monoidal structure on the topos.
    Examples include the Schanuel topos, which was explored extensively in \cite{schopp},
    and the classifying topos of noncommutative rings, which yields some sort of noncommutative synthetic differential geometry.
\end{abstract}

\listoftodos
\tableofcontents

\section{Introduction}

Every topos comes with an associated \emph{internal language}.
Each is built on intuitionistic logic, yet each has its own nature.
These logics can be studied for their own sake (e.g. \cite{sdg}),
or to learn more about the toposes they're built from (e.g. \cite{blechschmidt-thesis}).

Sometimes, the topos we're interested has an important monoidal product, distinct from both the product and coproduct.
So we would like the internal logic to likewise support this monoidal product.
While we could simply add axioms to the logic, saying this monoidal structure exists,
it's much more natural to move to a logical system that has the monoidal structure built in.
Such a thing is called \emph{bunched logic}.

In particular, \cite{schopp} introduces a bunched dependent type theory, which has the primitives we need.
But \cite{schopp} only applies their type theory to one particular case: the Schanuel topos.
If we want to apply it more generally, we need to prove a number of properties of our topos.

I streamline this process. Specializing to classifying toposes,
this paper provides an easy-to-check condition on a theory \(\T\)
that suffices to interpret \cite{schopp}'s type theory
in the classifying topos \(\Set[\T]\).

\section{Main Result}

\subsection{Setup}

Let \(\T\) be a geometric theory.

Suppose that for each pair \(A,B\) of sorts of \(\T\),
we have a relation \(\sim : A \times B \to \mathbf{Prop}\), geometrically defined in terms of \(\T\)'s signature.
For instance, if \(\T\) is the theory of rings, then \(\sim : R \times R \to \mathbf{Prop}\)
might be defined as \(a \sim b \iff ab = ba\).

For each sort \(A\), define the shorthand \(A_{\sim x} = \{a : A \mid a \sim x\}\).

We want to require these subsets to themselves form a \(\T\)-model.
If \(\T\) is the theory of rings, for instance, we want \(R_{\sim x}\) to be a ring, for any \(x : R\).

To state this precisely, say that \(\sim\) \emph{cuts out a sub-\(\T\)-model} if the following two conditions hold.
\begin{itemize}
    \item For any function symbol \(f : A_1 \times \dots \times A_n \to B\) and any \(x : X\),
    the restriction \(f_{\sim x} : (A_1)_{\sim x} \times \dots \times (A_n)_{\sim x} \to B_{\sim x}\) is well-defined for all \(x\).
    \item For any \(x : X\), every axiom of \(\T\) continues to hold when each sort \(A\) is replaced by \(A_{\sim x}\), each function symbol \(f\) is replaced by \(f_{\sim x}\), and each relation symbol is replaced by its restriction.
\end{itemize}

Let's look at the example where \(\T\) is the theory of \emph{local} rings,
where we define \(a \sim b \iff ab = ba\).

The first condition expands to the following five sequents.
\begin{align*}
    x : R, a : R, ax = xa, b : R, bx = xb &\vdash (a+b)x = x(a+b)
    \\ x : R, a : R, ax = xa, b : R, bx = xb &\vdash (ab)x = x(ab)
    \\ x : R, a : R, ax = xa &\vdash (-a)x = x(-a)
    \\ x : R &\vdash 0x = x0
    \\ x : R &\vdash 1x = x1
\end{align*}
In other words, it says that \(R_{\sim x} = \{a : R \mid a \sim x\}\) is a subring of \(R\), for all \(x : R\).

The second condition generates one requirement per axiom in the theory, but most of these are trivial.
The interesting one comes from the following axiom of local rings.
\begin{align*}
    a : R, b : R, (\exists c : R, (a+b)c = 1) &\vdash (\exists c : R, ac = 1) \lor (\exists c : R, bc = 1)
\end{align*}

The requirement generated is as follows.
\begin{align*}
    &x : R, a : R, ax = xa, b : R, bx = xb, (\exists c : R, cx = xc \land (a+b)c = 1)
    \\ &\vdash (\exists c : R, cx = xc \land ac = 1) \lor (\exists c : R, cx = xc \land bc = 1)
\end{align*}
This is exactly the condition that the subring \(R_{\sim x} \subseteq R\) is local.

\subsection{Theorem}

Without further ado, here is the main theorem of this paper.

\begin{theorem} \label{main}
    Let \(\T\) be a geometric theory.
    For each pair of sorts \(A,B\) of \(\T\), let \(\sim : A \times B \to \mathbf{Prop}\) be a geometrically-defined relation which cuts out a sub-\(\T\)-model.
    Further suppose \(\sim\) is symmetric; we have \(a : A, b : B, a \sim b \vdash b \sim a\) for any sorts \(A\) and \(B\).
    Then we can interpret the type theory \(\mathbf{BT}(*, 1, \Sigma, \Pi, \Pi^*)\), as defined in \cite[Section 5.1]{schopp}, in \(\Set[\T]\).
\end{theorem}

\subsection{Examples}

\begin{enumerate}
    \item 
    Consider the theory of rings, with \(a \sim b \iff ab = ba\).
    This is clearly symmetric and geometrically defined, and it cuts out a submodel
    because \(\{a : R \mid ax = xa\}\) is a subring of \(R\) for any \(x : R\).

    So Theorem \ref{main} applies, and we can interpret \(\mathbf{BT}(*, 1, \Sigma, \Pi, \Pi^*)\) in \(\Set[\Ring]\).

    This is the example that motivated me to write this paper.

    \item
    Let \(\T\) be the theory of infinite decidable objects.
    Its classifying topos is known as the Schanuel topos.
    \begin{align*}
        &N \text{ sort}
        \\ a : N, b : N &\vdash a \mathrel{\#} b  \text{ prop}
        \\ a : N, a \mathrel{\#} a &\vdash \bot
        \\ a : N, b : N &\vdash a = b \lor a \mathrel{\#} b
        \\ a_1 : N, \dots, a_n : N &\vdash \exists b : N, \bigwedge_i a_i \mathrel{\#} b
    \end{align*}
    Choose \(\sim\) to be the apartness relation: \(a \sim b \iff a \mathrel{\#} b\).
    This is symmetric and geometrically defined. And since \(N\) is an infinite decidable object,
    so is \(\{a : N \mid a \mathrel{\#} x\}\).

    So we can interpret \(\mathbf{BT}(*, 1, \Sigma, \Pi, \Pi^*)\) in the Schanuel topos.
    This is the example discussed in \cite{schopp}.

    \item
    For any theory, there is a trivial example where \(a \sim b\) always holds.
    So we can interpret \(\mathbf{BT}(*, 1, \Sigma, \Pi, \Pi^*)\) in any topos.
    This turns out to be the special case where the monoidal product coincides with the cartesian product.

    \item There seems to be a ``generic example'' in the theory of undirected graphs.
    \begin{align*}
        &V \text{ sort}
        \\ a : V, b : V &\vdash a \sim b \text{ prop}
        \\ a : V, b : V, a \sim b &\vdash b \sim a
    \end{align*}
\end{enumerate}

\section{Proof}

We begin by bringing in some results from \cite{schopp}.

\begin{lemma}
    The type theory \(\mathbf{BT}(*, 1, \Sigma, \Pi, \Pi^*)\) can be modeled in any \\
    \((*, 1, \Sigma, \Pi, \Pi^*)\)-type-category.
\end{lemma}

\begin{proof}
    \cite{schopp}, section 6.
\end{proof}

\begin{lemma}
    Let \(\B\) be a topos equipped with a strict affine symmetric monoidal closed structure \((*,\wand)\),
    such that \(- * A\) preserves pullbacks for all \(A\).
    Further suppose we have a canonical choice of pullback for every span in \(\B\).
    Then the family fibration over \(\B\) is a \((*, 1, \Sigma, \Pi, \Pi^*)\)-type-category.
\end{lemma}

\begin{proof}
    \cite{schopp}, section 3.5. Since we are in a topos, the condition that the monomorphism \(A * B \hookrightarrow A \times B\) is strong is trivial.
\end{proof}

Next, we specialize to sheaf toposes.

\begin{lemma}
    Let \((C,J)\) be a (small) site, such that \(C\) is finitely complete.
    In fact, suppose we have a canonical choice for pullbacks in \(C\).
    Let \(*\) be an affine symmetric monoidal structure on \(C\) that preserves pullbacks and covers.
    Then \(\Sh(C,J)\) satisfies the conditions of the above lemma.
\end{lemma}

\begin{proof}
    \(\Sh(C,J)\) is a Grothendieck topos, hence a topos.

    \vspace*{1em}\hrule\vspace*{1em}

    Theorem 4.3.2 from \cite{biering} says we get a monoidal structure \(\otimes^\Sh\) on \(\Sh(C,J)\)
    by transporting Day convolution across the adjunction \((\mathbf{a} \dashv i) : \Psh(C,J) \rightleftarrows \Sh(C,J)\).
    Its monoidal unit is the sheafification of the Yoneda embedding of \(*\)'s unit, which simplifies to the terminal object.
    Chasing definitions quickly shows \(\otimes^\Sh\) is symmetric.
    And corollary 4.3.10 from \cite{biering} says this monoidal structure is closed.

    In other words, \(\otimes^\Sh\) is an affine symmetric monoidal closed structure.
    We must show this is strict --- that the canonical map \(A \otimes^\Sh B \to A \times B\) is always a monomorphism.

    But, letting \(\otimes^\Psh\) represent Day convolution on \(\Psh(C)\),
    this map is just \(iA \otimes^\Psh iB \to iA \times iB\), restricted to sheaves.
    Expanding both \(\otimes^\Psh\) and \(\times\) as Day convolutions,
    this is a map, natural in \(c \in C^\mathrm{op}\), of the following type.
    \begin{scriptsize}\[\left(\int^{c_1,c_2 \in C} A(c_1) \times B(c_2) \times (c \xrightarrow{C} c_1 * c_2)\right) \to \left(\int^{c_1,c_2 \in C} A(c_1) \times B(c_2) \times (c \xrightarrow{C} c_1 \times c_2)\right)\]\end{scriptsize}

    This map is exactly what you'd expect it to be; it composes the \(c \xrightarrow{C} c_1 * c_2\) component
    with the canonical map \(c_1 * c_2 \xrightarrow{C} c_1 \times c_2\), and leaves everything else alone.
    So since that canonical map is mono, this is too.

    \vspace*{1em}\hrule\vspace*{1em}

    Finally, we have a canonical construction for pullbacks in \(\Sh(C,J)\), since we have one for \(C\).
\end{proof}

And finally, we set the site to a syntactic site, to understand classifying toposes.

\begin{lemma}
    Let \(\T\) be a geometric theory.
    For each pair of sorts \(A,B\) of \(\T\), let \(\sim : A \times B \to \mathbf{Prop}\) be a symmetric, geometrically-defined relation, which cuts out a sub-\(\T\)-model.
    Then the syntactic site for \(\T\) satisfies the conditions of the above lemma.
\end{lemma}

\begin{proof}
    Let us begin with a few shorthands.

    We'll allow ourselves to write sequents involving compound types, such as \(p : A \times B \vdash \pi_1 p \sim \pi_2 p\).
    These can be straightforwardly ``compiled out'' to ordinary sequents, such as \(a : A, b : B \vdash a \sim b\).

    Next, if we have a context \(\Gamma = (a_1 : A_1, \dots, a_n : A_n, \phi_1 \dots \phi_p)\),
    we'll reuse the name \(\Gamma\) for the type \(\{(a_1, \dots, a_n) : A_1 \times \dots \times A_n \mid \phi_1 \land \dots \land \phi_p\}\).
    If we have a substitution \(f : \Gamma \to \Delta\), there is then a natural way to define the term \(\gamma : \Gamma \vdash f\gamma : \Delta\).

    Finally, if \(\gamma = (a_1, \dots, a_n) : \Gamma\) and \(\delta = (b_1, \dots, b_m) : \Delta\),
    we write \(\gamma \sim \delta\) as a shorthand for \(\bigwedge_{i=1}^m \bigwedge_{j=1}^n a_i \sim b_j\).

    With that out of the way, let's begin the proof.

    \vspace*{1em}\hrule\vspace*{1em}

    First of all, the syntactic category is guaranteed to be finitely complete, with a canonical construction for pullbacks.
    Specifically, the empty context is terminal,
    and the pullback of the span \(\Gamma \xrightarrow{f} \Xi \xleftarrow{g} \Delta\) is the context \((\gamma : \Gamma, \delta : \Delta, f\gamma = g\delta)\).

    Second, we describe the monoidal structure. Given contexts \(\Gamma\) and \(\Delta\),
    define \(\Gamma * \Delta = (\gamma : \Gamma, \delta : \Delta, \gamma \sim \delta)\).

    Functoriality of \(*\) reduces to the fact that \(\sim\) respects substitutions, which further reduces to the fact that \(\sim\) respects function symbols.
    Symmetry reduces to the symmetry of \(\sim\). Choosing the empty context as monoidal unit, the unit laws hold on the nose, and associativity up to a rearrangement of contexts.
    All the coherences hold.
    And since \(\Gamma * \Delta\) is just \(\Gamma \times \Delta\) with extra propositional hypotheses,
    the map \(\Gamma * \Delta \hookrightarrow \Gamma \times \Delta\) is a monomorphism.
    Putting that all together, \(*\) is a strict affine symmetric monoidal product on the syntactic category.

    Third, we show that starring preserves pullbacks.

    Say we have a span \(\Gamma \xrightarrow{f} \Xi \xleftarrow{g} \Delta\), and another context \(\Theta\).
    The pullback of the span is the context \((\gamma : \Gamma, \delta : \Delta, f\gamma = g\delta)\).
    Starring with \(\Theta\) yields \((\gamma : \Gamma, \delta : \Delta, f\gamma = g\delta, \theta : \Theta, \gamma \sim \theta, \delta \sim \theta)\).

    On the other hand, if we star with \(\Theta\) first, we get the span
    \((\gamma : \Gamma, \theta : \Theta, \gamma \sim \theta) \xrightarrow{f * \mathrm{id}_\Theta} (\xi : \Xi, \theta : \Theta, \xi \sim \theta) \xleftarrow{g * \mathrm{id}_\Theta} (\delta : \Delta, \theta : \Theta, \delta \sim \theta)\).
    Taking the pullback yields \((\gamma : \Gamma, \theta_1 : \Theta, \gamma \sim \theta_1, \delta : \Delta, \theta_2 : \Theta, \delta \sim \theta_2, f\gamma = g\delta, \theta_1 = \theta_2)\),
    which is equivalent to what we got before.

    And fourth, we show that starring preserves covers.

    Suppose \(\{f_i : \Gamma_i \to \Delta \mid i \in I\} \in J\).
    This means \(\delta : \Delta \vdash \bigvee_{i \in I} \exists \gamma : \Gamma_i, f_i\gamma = \delta\) is provable.

    By the final hypothesis, the proof is still valid if we, throughout the proof, replace each sort \(X\) with \(X_{\sim \theta} = \{x : X \mid x \sim \theta\}\), and each function \(f\) with its restriction \(f_{\sim \theta}\).
    This is only stated when \(\theta : \Theta\) for some \emph{sort} \(\Theta\),
    but the same is true for any \emph{context} \(\Theta\),
    simply by iterating once for each variable in \(\Theta\).

    So \(\theta : \Theta, \delta : \Delta_{\sim \theta} \vdash \bigvee_{i \in I} \exists \gamma : (\Gamma_i)_{\sim \theta}, f_i\gamma = \delta\).
    But this is equivalent to the statement \(\delta : (\Delta * \Theta) \vdash \bigvee_{i \in I} \exists \gamma : (\Gamma_i * \Theta), (f_i * \mathrm{id}_\Theta)\gamma = \delta\),
    so \(\{f_i * \mathrm{id}_\Theta : \Gamma_i * \Theta \to \Delta * \Theta \mid i \in I\}\) is a cover, as required.
\end{proof}

Putting this all together:

\begin{proof}[Proof of Theorem \ref{main}]
    Combine the above four lemmas.
\end{proof}

\section{Using the Monoidal Structure}

It's not enough to interpret the type theory in our topos of choice.
We need some axioms to build on, telling us how the new monoidal structure behaves.
After all, there's no rule in the type theory to say \(*\) isn't the same thing as \(\times\)!

\begin{theorem}
    Recall that \(\Set[\T]\) contains the generic model of \(T\).
    So for any sort \(X\) in \(\T\), let \(U_X\) be the object of \(\Set[\T]\) that models it.
    And for any sorts \(X\) and \(Y\), let \(\sim_U : U_X \times U_Y \to \mathbf{Prop}\) be the internal relation modelling \(\sim\).
    Then, in the internal logic, the image of the inclusion \(U_X * U_Y \hookrightarrow U_X \times U_Y\)
    is \(\{(x,y) \mid x \sim_U y\}\).
\end{theorem}

In other words, in the Schanuel topos, \(N * N \simeq \{(x,y) : N^2 \mid x \mathrel{\#} y\}\).
In \(\Set[\Ring]\), \(R * R \simeq \{(x,y) : R^2 \mid xy = yx\}\).
% In \(\Set[\Graph]\), \(V * V\) is the set of edges of the generic graph. issue: duplicated

\begin{proof}
    \(U_X\) is given by the sheafification of the Yoneda embedding of the context containing a single variable of type \(x\).
    In short, \(U_X = \mathbf{a}\yo X\). Similarly for \(U_Y\). So we calculate, using Lemma 4.3.1 from \cite{biering}:
    \begin{align*}
        U_X * U_Y &= \mathbf{a}\yo X * \mathbf{a}\yo Y
        \\ &= \mathbf{a}(i\mathbf{a}\yo X \otimes_\mathrm{Day} i\mathbf{a}\yo Y)
        \\ &= \mathbf{a}(\yo X \otimes_\mathrm{Day} i\mathbf{a}\yo Y)
        \\ &= \mathbf{a}(\yo X \otimes_\mathrm{Day} \yo Y)
        \\ &= \mathbf{a}(c \mapsto \int^{c_1,c_2} (\yo X)c_1 \times (\yo Y)c_2 \times (c \xrightarrow{C} c_1 * c_2))
        \\ &= \mathbf{a}(c \mapsto \int^{c_1,c_2} (c_1 \xrightarrow{C} X) \times (c_2 \xrightarrow{C} Y) \times (c \xrightarrow{C} c_1 * c_2))
        \\ &= \mathbf{a}(c \mapsto (c \xrightarrow{C} X * Y))
        \\ &= \mathbf{a}(\yo (X * Y))
    \end{align*}

    This is the internalization of the context \(X * Y = (x : X, y : Y, x \sim y)\),
    so it equals \(\{(x : U_X, y : U_Y) \mid x \sim_U y\}\).
\end{proof}

This shows what happens when we monoidally combine types taken from the generic model.
It's also interesting consider the opposite. What happens if we combine types that don't mention the generic model at all?
\begin{theorem}
    Let \(A, B \in \Set[\T]\), where \(A\) is the inverse image of a set \(\mathcal{A}\) under the global sections functor.
    Then \(A * B \simeq A \times B\).
\end{theorem}
So, for example, \(\mathbf{2} * X \simeq \mathbf{2} \times X \simeq X + X\).

\begin{proof}
    We have the following chain of isomorphisms:
    \begin{small}
    % https://q.uiver.app/#q=WzAsOSxbMCwwLCJcXGhvbShBICogQiwgWCkiXSxbMCwxLCJcXGhvbShBLCBCIFxcd2FuZCBYKSJdLFswLDIsIlxcaG9tXFxsZWZ0KFxcbGVmdChcXGNvcHJvZFxcbGltaXRzX3thIFxcaW4gXFxtYXRoY2Fse0F9fSBcXG1hdGhiZnsxfVxccmlnaHQpLCBCIFxcd2FuZCBYXFxyaWdodCkiXSxbMCwzLCJcXHByb2RcXGxpbWl0c197YSBcXGluIFxcbWF0aGNhbHtBfX0gXFxob20oXFxtYXRoYmZ7MX0sIEIgXFx3YW5kIFgpIl0sWzEsMywiXFxwcm9kXFxsaW1pdHNfe2EgXFxpbiBcXG1hdGhjYWx7QX19IFxcaG9tKEIsIFgpIl0sWzIsMywiXFxwcm9kXFxsaW1pdHNfe2EgXFxpbiBcXG1hdGhjYWx7QX19IFxcaG9tKFxcbWF0aGJmezF9LCBCIFxcdG8gWCkiXSxbMiwyLCJcXGhvbVxcbGVmdChcXGxlZnQoXFxjb3Byb2RcXGxpbWl0c197YSBcXGluIFxcbWF0aGNhbHtBfX0gXFxtYXRoYmZ7MX1cXHJpZ2h0KSwgQiBcXHRvIFhcXHJpZ2h0KSJdLFsyLDEsIlxcaG9tKEEsIEIgXFx0byBYKSJdLFsyLDAsIlxcaG9tKEEgXFx0aW1lcyBCLCBYKSJdLFswLDEsIiIsMCx7InN0eWxlIjp7InRhaWwiOnsibmFtZSI6ImFycm93aGVhZCJ9fX1dLFsxLDIsIiIsMCx7InN0eWxlIjp7InRhaWwiOnsibmFtZSI6ImFycm93aGVhZCJ9fX1dLFsyLDMsIiIsMCx7InN0eWxlIjp7InRhaWwiOnsibmFtZSI6ImFycm93aGVhZCJ9fX1dLFszLDQsIiIsMCx7InN0eWxlIjp7InRhaWwiOnsibmFtZSI6ImFycm93aGVhZCJ9fX1dLFs0LDUsIiIsMCx7InN0eWxlIjp7InRhaWwiOnsibmFtZSI6ImFycm93aGVhZCJ9fX1dLFs1LDYsIiIsMCx7InN0eWxlIjp7InRhaWwiOnsibmFtZSI6ImFycm93aGVhZCJ9fX1dLFs2LDcsIiIsMCx7InN0eWxlIjp7InRhaWwiOnsibmFtZSI6ImFycm93aGVhZCJ9fX1dLFs3LDgsIiIsMCx7InN0eWxlIjp7InRhaWwiOnsibmFtZSI6ImFycm93aGVhZCJ9fX1dXQ==
    \[\begin{tikzcd}
        {\hom(A * B, X)} && {\hom(A \times B, X)} \\
        {\hom(A, B \wand X)} && {\hom(A, B \to X)} \\
        {\hom\left(\left(\coprod\limits_{a \in \mathcal{A}} \mathbf{1}\right), B \wand X\right)} && {\hom\left(\left(\coprod\limits_{a \in \mathcal{A}} \mathbf{1}\right), B \to X\right)} \\
        {\prod\limits_{a \in \mathcal{A}} \hom(\mathbf{1}, B \wand X)} & {\prod\limits_{a \in \mathcal{A}} \hom(B, X)} & {\prod\limits_{a \in \mathcal{A}} \hom(\mathbf{1}, B \to X)}
        \arrow[tail reversed, from=1-1, to=2-1]
        \arrow[tail reversed, from=2-1, to=3-1]
        \arrow[tail reversed, from=2-3, to=1-3]
        \arrow[tail reversed, from=3-1, to=4-1]
        \arrow[tail reversed, from=3-3, to=2-3]
        \arrow[tail reversed, from=4-1, to=4-2]
        \arrow[tail reversed, from=4-2, to=4-3]
        \arrow[tail reversed, from=4-3, to=3-3]
    \end{tikzcd}\]
    \end{small}

    So the result follows by Yoneda.
\end{proof}

\section{Applications}

The main purpose of this paper is to enable the exploration of monoidal toposes via their internal logic.

\subsection{Schanuel Topos, and Variations} \label{section-schanuel-and-variations}

\cite{schopp} does a good job of this for the Schanuel topos, the classifying topos of infinite decidable objects.
They focus on modeling the syntax of programming languages, using the generic infinite decidable object as their type of abstract variables.

Our formulation suggests an interesting extension to this idea.
Instead of infinite decidable \emph{objects}, with \(\sim\) being apartness,
what if we work with unbounded dense decidable \emph{linear orders}, with \(\sim\) being the strict inequality?
We lose symmetry, so we'd need to modify the type theory to make \(*\) asymmetric.
But now our abstract variables come with a natural ordering relation, where \(A * B\)
means all the variables in \(A\) come \emph{before} all the variables in \(B\).

Is this useful? I have no idea. Is it interesting? Definitely!

\subsection{Noncommutative Synthetic Differential Geometry}

I was inspired to write this paper by a use-case related to \emph{synthetic differential geometry}.

Synthetic differential geometry (\cite{sdg}) is the study of toposes whose internal logics satisfy, at minimum, the following axioms.
\begin{itemize}
    \item There is a commutative ring, \(R\).
    \item Let \(D = \{x \in R | x^2 = 0\}\). Then the map \(R^2 \to (D \to R)\) given by \((a,b) \mapsto (d \mapsto a + bd)\) is an isomorphism.
\end{itemize}

This allows the immediate definition of derivatives, tangent vectors, or other concepts from differential geometry.
For instance, if \(f : R \to R\), \(f'\) is simply the unique function satisfying \(f(x+d) = f(x) + f'(x) d\) for all \(x : X\) and \(d : D\).
Limits and complicated epsilon-delta arguments are completely sidestepped; \emph{every} function \(R \to R\) is smooth.

A prototypical, though suboptimal, topos satisfying these axioms is the classifying topos of commutative rings.
But this immediately suggests a \emph{noncommutative} version of synthetic differential geometry, where we work with \(\Set[\Ring]\) instead of \(\Set[\CRing]\).

When I tried this, the first thing I discovered is that \(\Set[\Ring]\) has an important monoidal structure,
given by Day convolution over the tensor product of rings.
This is exactly the monoidal structure we get if we set \(a \sim b = (ab = ba)\) in Theorem \ref{main}.

And indeed, \(\Set[\Ring]\) supports a \emph{monoidal} version of the synthetic differential geometry axioms.
\begin{theorem} \label{nsdg}
    Internally to \(\Set[\Ring]\), let \(R\) be the generic ring, and let \(D = \{x \in R | x^2 = 0\}\).
    Then \(R^2 \simeq (D \wand R)\).
\end{theorem}

\begin{proof}
    We reason analogously to the proof in section 9.3 of \cite{sgm}.

    Recall that \(\Set[\Ring]\) is the topos of presheaves on the opposite of the category of finitely presented rings.
    We have \(R = \yo(\Z[x])\), \(D = \yo(\Z[d]/(d^2))\), and \(R^2 = \yo(\Z[a,b])\).

    We want to show \(R^2\) is the monoidal exponential \(D \wand R\).
    So since monoidal exponentials are preserved by Yoneda,
    it suffices to show that \(\Z[a,b]\) is a monoidal exponential of \(\Z[x]\) and \(\Z[d]/(d^2)\) in the opposite of the category of finitely presented rings.
    That is, that it's the monoidal \emph{coexponential} in finitely presented rings.

    We check the universal property directly.

    Define the coevaluation map \(\epsilon \in \hom_{\Ring_\mathrm{fp}}(\Z[x], \Z[a,b] \otimes \Z[d]/(d^2))\)
    to be the unique map taking \(x\) to \(a \otimes 1 + b \otimes d\).

    Now, every \(f \in \hom_{\Ring_\mathrm{fp}}(\Z[x], X \otimes \Z[d]/(d^2))\) is determined by where it maps \(x\).
    And \(f(x)\) will always have the form \(\alpha \otimes 1 + \beta \otimes d\), for some unique choice of \(\alpha, \beta \in X\).
    But then \(f = (g \otimes \mathrm{id}) \circ \epsilon\),
    where \(g\) maps \(a\) to \(\alpha\) and \(b\) to \(\beta\).

    So every \(f \in \hom_{\Ring_\mathrm{fp}}(\Z[x], X \otimes \Z[d]/(d^2))\) factors through \(\epsilon\).
    And this factorization is unique, because every possible \(g\) is hit by the procedure just described.
\end{proof}

A final point of interest: Internally to \(\Set[\CRing]\), the type \(R \to R\) is exactly the space of polynomials.
But when we move to \(\Set[\Ring]\) and \(R\) becomes noncommutative, there are now \emph{two} reasonable notions of polynomial.
Either the coefficients are assumed to commute with the indeterminate, or they aren't.
These two notions correspond exactly to the two choices of function space: \(R \wand R\) and \(R \to R\).

\subsection{Separation Logic}

A more speculative application is \emph{separation logic},
a method of proving the correctness of imperative programs.

Traditionally, separation logic is an intuitionistic logic, equipped with an extra monoidal closed structure.
To me, this looks like a \emph{decategorified} monoidal topos.
Might there be some clever choice of theory \(T\) and relation \(\sim\),
so that the interpretation of \(\mathbf{BT}(*, 1, \Sigma, \Pi, \Pi^*)\)
in the classifying topos \(\Set[\T]\) acts as a categorified separation logic?

\section{Future Work}

Besides the applications detailed above, I can see a few other worthwhile lines of research.

\subsection{Reasoning Principles}

When discussing the application to \(\Set[\Ring]\), the proof of Theorem \ref{nsdg}
stepped outside the internal logic, reasoning directly about the topos in its incarnation as a functor category.
I find this unsatisfying.

By contrast, when proving the version of this for \(\Set[\CRing]\),
we can begin with the axiom scheme provided by \cite{blechschmidt},
then proceed to work entirely within the internal logic.
Specifically, the axiom tells us that \(A \simeq (\hom_{R\text{-alg}}(A,R) \to R)\) for any \(R\)-algebra \(A\);
specializing to \(A = R[d]/(d^2)\) shows that \((D \to R) \simeq R[d]/(d^2)\), which is set-isomorphic to \(R^2\).

It would be valuable to either extend \cite{blechschmidt}'s axiom scheme to allow internal proofs of theorems like Theorem \ref{nsdg},
or show how to do so using the existing axiom scheme.

\subsection{Modifying the Type Theory}

It might also make sense to explore modifications or extensions to the type theory.

We saw this already in Section \ref{section-schanuel-and-variations},
where the theory of unbounded dense decidable linear orders wants \(*\) to be asymmetric.

It also seems that the \emph{relation} \(\sim : A \times B \to \mathbf{Prop}\) ought to be generalized to a \emph{dependent type} \(\sim : A \times B \to \mathbf{Type}\).
% and that this would correspond to the internal maps \(A * B \to A \times B\) no longer being injective.
What changes to \cite{schopp}'s type theory does this require, if any?

And finally, \cite{schopp} found it useful to add a monoidal \(\Sigma\) type to their type theory, analogous to the monoidal \(\Pi\) type.
What additional structure is needed on \(\T\) for the interpretation in \(\Set[\T]\) to support this?

\todo{References}
\todo{General proofreading}

\end{document}

### 2025-06-11 paper.tex
minor rephrase

@@
 
-\cite{schopp} introduces a bunched dependent type theory, \(\mathbf{BT}(*, 1, \Sigma, \Pi, \Pi^*)\),
-of potentially general utility. But they only instantiate it for a single topos: the Schanuel topos.
+Dependent type theory, in its standard form, has a model in every topos.
+By working within the type theory, you can therefore reason ``internally'' to the topos.
 
-By stitching together a few results from the literature,
+To capture an additional monoidal structure on the topos,
+\cite{schopp} extends the type theory to a system they call \(\mathbf{BT}(*, 1, \Sigma, \Pi, \Pi^*)\).
+While that paper is specific to the Schanuel topos, the type theory described within is applicable to many others.
+
 I provide an easy method to verify that \(\mathbf{BT}(*, 1, \Sigma, \Pi, \Pi^*)\)
 can be interpreted in a classifying topos of your choice,
-so you can start \emph{using} the language as quickly as possible.
+so you can start \emph{using} the internal language as quickly as possible.
 

### 2025-06-26 paper.tex
Fix issue #1.

@@
 \usepackage{mathpartir}
-\usepackage{tcolorbox}
 
@@
 \newtheorem{lemma}[theorem]{Lemma}
+\newtheorem{construction}[theorem]{Construction}
 \newtheorem{example}[theorem]{Example}
@@
 
+For simplicity of exposition, I describe the theorem and proof in the case of single-sorted theories.
+The extension to multiple sorts is straightforward, but terminologically awkward.
+
 \begin{theorem}
-    Consider a geometric theory, \(\T\).
-    For each pair \(A,B\) of sorts of \(\T\), let \((\sim) : A \times B \to \Prop\) be a relation, defined in terms of \(\T\)'s signature.
-    Further suppose that \((\sim)\) is symmetric, and that the following holds.
-
-    \begin{quote}
-        Let \(X\) be an arbitrary sort of \(\T\), and let \(x : X\).
-        Then if you restrict each sort \(A\) of \(\T\) to the subset \(\{a : A \mid a \sim x\}\),
-        the result is another \(\T\)-model.
-    \end{quote}
-
-    \noindent Then \(\mathbf{BT}(*, 1, \Sigma, \Pi, \Pi^*)\) interprets into \(\Set[\T]\).
+    Let \(\T\) be a single-sorted geometric theory, with sort \(X\).
+    Let \((\sim) : X \to X \to \Prop\) be a geometrically-defined symmetric relation.
+    Further suppose that for any \(x \in X\), \(\{a : X \mid a \sim x\}\) is a sub-\(\T\)-model of \(X\).
+    Then \(\mathbf{BT}(*, 1, \Sigma, \Pi, \Pi^*)\) interprets into \(\Set[\T]\).
 \end{theorem}
 
-\subsection{Example}
-
-As an example, let \(\T\) be the theory of local rings, and let \(\sim : R \times R \to \Prop\) be the relation \(a \sim b \iff a \cdot b = b \cdot a\).
-
-\((\sim)\) is clearly symmetric. And for any \(x : R\), \(\{a : R \mid a \sim x\}\) is a local subring of \(R\).
-So the conditions of the theorem are satisfied, and therefore the type theory interprets into \(\Set[\T]\).
+For example, let \(\T\) be the theory of rings, and let \(a \sim b\) mean \(ab = ba\).
+The hypothesis holds because centralizers of rings are rings;
+so we get an interpretation of the bunched type theory into \(\Set[\Ring]\).
 
@@
 
-Some of the concepts in the theorem statement may be insufficiently precise. I address that here.
-
-To say \(\sim\) is symmetric is to say that for every pair \(A,B\) of sorts in \(\T\),
-the following is provable in the language of the theory.
-\[a : A, b : B \vdash a \sim b \implies b \sim a\]
-
-To say that the subsets \(\{a : A \mid a \sim x\}\) form another \(\T\)-model is to say the following:
+The above makes more sense intuitively than formally.
+So to be perfectly precise, when I say
+``\(\{a : X \mid a \sim x\}\) is a sub-\(\T\)-model of \(X\)'',
+I mean the following:
 \begin{itemize}
-    \item For every sort \(X\) and every function symbol \(f : A_1 \times \dots \times A_n \to B\), the following is provable:
-    \[x : X, a_1 : A_1 \dots a_n : A_n, a_1 \sim x \dots a_n \sim x \vdash f(a_1 \dots a_n) \sim x\]
-    \item For every sort \(X\) and every axiom \(a_1 : A_1 \dots a_n : A_n, P \vdash Q\):
-
-    Let \(P'\) be produced from \(P\)
-    by replacing each existential quantifier \(\exists b : B, \dots\)
-    with \(\exists b : B, b \sim x \land \dots\).
-    Similarly define \(Q'\) in terms of \(Q\).
-    Then the following is provable:
-    \[x : X, a_1 : A_1 \dots a_n : A_n, a_1 \sim x \dots a_n \sim x, P' \vdash Q'\]
+    \item For every function symbol \(f : X^n \to X\), we require:
+    \[x : X, \substack{a_1 : X \\ a_1 \sim x}, \dots, \substack{a_n : X \\ a_n \sim x} \vdash f(a_1 \dots a_n) \sim x\]
+    \item For every axiom of \(\T\), run the following steps. We require the resulting sequent to hold.
+    \begin{itemize}
+        \item Prepend a variable \(x : X\) to the context.
+        \item For each other variable \(a : X\) in the context, add an additional hypothesis \(a \sim x\).
+        \item Replace every instance of \(\exists a : X, P\) with \(\exists a : X, a \sim x \land P\).
+    \end{itemize}
 \end{itemize}
 
-To see how this plays out in practice, consider the example from above --- the theory of local rings.
-
-\begin{tcolorbox}[colback=red!5!white, colframe=red!50!black]
-    \begin{mathpar}
-        R \text{ sort}
-        \and 0 : R
-        \and 1 : R
-        \\ (+) : R \times R \to R
-        \and (\cdot) : R \times R \to R
-        \and (-) : R \to R
-        \\ a,b,c : R \vdash (a + b) + c = a + (b + c)
-        \and a,b : R \vdash a + b = b + a
-        \\ a : R \vdash a + 0 = a
-        \and a : R \vdash a + (-a) = 0
-        \\ a,b,c : R \vdash (a \cdot b) \cdot c = a \cdot (b \cdot c)
-        \and a : R \vdash a \cdot 1 = a
-        \\ a,b,c : R \vdash (a + b) \cdot c = a \cdot c + b \cdot c
-        \and a,b,c : R \vdash a \cdot (b + c) = a \cdot b + a \cdot c
-        \\ 0 = 1 \vdash \bot
-        \and a,b : R, a + b = 1 \vdash
-            (\exists c, a \cdot c = 1 \land c \cdot a = 1) \lor
-            (\exists c, b \cdot c = 1 \land c \cdot b = 1)
-    \end{mathpar}
-    \tcblower
-    \[a \sim b \iff a \cdot b = b \cdot a\]
-\end{tcolorbox}
-
-The sub-\(\T\)-model conditions say the following:
-\begin{tcolorbox}[colback=blue!5!white, colframe=blue!50!black]
-    \begin{align*}
-        x : R &\vdash 0 x = x 0
-        \\ x : R &\vdash 1 x = x 1
-        \\ x,a,b : R, a x = x a, b x = x b &\vdash (a + b) x = x (a + b)
-        \\ x,a,b : R, a x = x a, b x = x b &\vdash (a b) x = x (a b)
-        \\ x,a : R, a x = x a &\vdash (-a) x = x (-a)
-    \end{align*}
-    \tcblower
-    \begin{align*}
-        x,a,b,c : R, ax=xa, bx=xb, cx=xc &\vdash (a + b) + c = a + (b + c)
-        \\ x,a,b : R, ax=xa, bx=xb &\vdash a + b = b + a
-        \\ x,a : R, ax=xa &\vdash a + 0 = a
-        \\ x,a : R, ax=xa &\vdash a + (-a) = 0
-        \\ x,a,b,c, ax=xa, bx=xb, cx=xc : R &\vdash (ab)c = a(bc)
-        \\ x,a : R, ax=xa &\vdash a1 = a
-        \\ x,a,b,c, ax=xa, bx=xb, cx=xc : R &\vdash (a + b) c = a c + b c
-        \\ x,a,b,c, ax=xa, bx=xb, cx=xc : R &\vdash a (b + c) = a b + a c
-        \\ x : R, 0 = 1 &\vdash \bot
-        \\ x,a,b : R, ax=xa, bx=xb, a + b = 1 &\vdash \\
-            (\exists c, cx=xc \land ac = 1 \land ca = 1) &\lor
-            (\exists c, cx=xc \land bc = 1 \land cb = 1)        
-    \end{align*}
-\end{tcolorbox}
-
-All of these are easily verified. More to the point, they capture the idea
-that for all \(x : R\), the subset \(\{a : R \mid ax = xa\}\) is a local ring.
+As an example of the latter, say \(\T\) is the theory of inhabited sets, with axiom \(\vdash \exists a : X, \top\).
+Then the required sequent is \(x : X \vdash \exists a : X, a \sim x \land \top\).
 
@@
 
-To prove the main theorem, we begin by recalling a result from \cite{schopp}.
+For the duration of this section, we accept the assumptions of the main theorem.
 
-\begin{lemma}
-    Let \(\B\) be a topos equipped with a strict affine symmetric monoidal closed structure \((*,\wand)\),
-    such that \(- * A\) preserves pullbacks for all \(A\).
-    Further suppose we have a canonical choice of pullback for every span in \(\B\).
-    Then the type theory \(\mathbf{BT}(*, 1, \Sigma, \Pi, \Pi^*)\) interprets into \(\B\).
+\vspace{1em}
+
+\noindent To begin, notice a key detail of the transformation described above.
+\begin{lemma} \label{transform-lemma}
+    The transformation of sequents described in section \ref{Details},
+    used there to transform the axioms of \(\T\), preserves derivability.
+    Any provable sequent remains provable after the transformation.
 \end{lemma}
@@
 \begin{proof}
-    Chapter 6 of \cite{schopp} shows that we can interpret the type theory in any ``\((*, 1, \Sigma, \Pi, \Pi^*)\)-type-category''.
-    Section 3.5 tells us that \(\B\) is such a category.
+    First, prove that for any term \(t\) in any context, \(t \sim x\) is provable in the transformed context.
+    This is immediate by induction on the structure of \(t\) ---
+    for variables, the transformation added \(t \sim x\) to the context;
+    for compound terms, \(f(t_1 \dots t_n) \sim x\) follows from the statements \(t_i \sim x\).
+
+    Now the lemma follows from a second induction, this time on the structure of proofs.
+    There are a number of cases; all of them easy, so I'll provide just two examples.
+
+    Consider the introduction and elimination rules for \(\exists\).
+    \begin{mathpar}
+        \inferrule{\Gamma \vdash t : X \\ \Gamma \vdash P[t/a]}{\Gamma \vdash \exists a:X, P}
+        \and \inferrule{\Gamma \vdash \exists a:X, P \\ \Gamma, a:X, P \vdash Q}{\Gamma \vdash Q}
+    \end{mathpar}
+    After transformation, these become:
+    \begin{mathpar}
+        \inferrule{\Gamma' \vdash t : X \\ \Gamma' \vdash P'[t/a]}{\Gamma' \vdash \exists a:X, a \sim x \land P'}
+        \and \inferrule{\Gamma' \vdash \exists a:X, a \sim x \land P' \\ \Gamma', a:X, a \sim x, P' \vdash Q'}{\Gamma' \vdash Q'}
+    \end{mathpar}
+    Both of which are immediate; the first because we've already proven \(\Gamma' \vdash t \sim x\).
 \end{proof}
 
-Next, we specialize to sheaf toposes.
+Next, bootstrap to the syntactic site.
+\begin{construction}
+    We place a monoidal structure on \(\T\)'s syntactic category.
+\end{construction}
 
-\begin{lemma} \label{sh}
-    Let \((C,J)\) be a (small) site, such that \(C\) is finitely complete,
-    with a canonical choice of pullbacks.
-    Let \(*\) be a strict affine symmetric monoidal structure on \(C\) that preserves pullbacks and covers.
-    Then \(\Sh(C,J)\) satisfies the conditions of the above lemma.
+\begin{proof}
+    Given two contexts \(A\) and \(B\), define their product by extending \(A \times B\) with the hypotheses \(a \sim b\) for each variable \(a\) in the first context and each variable \(b\) in the second.
+    For example:
+    \[(a,b : X, \phi(a,b)) * (c : X, \psi(c)) = (a,b,c : X, \phi(a,b), \psi(c), a \sim c, b \sim c)\]
+
+    We'll handle functoriality one argument at a time.
+
+    Given a morphism \(f : A \to A'\), we must produce a morphism \(f*1_B : A*B \to A'*B\).
+    That is, a provably functional relation on the combined context \((A*B) \times (A'*B)\).
+
+    But \(f \times 1_B : A \times B \to A' \times B\) is similarly a relation on the context \((A \times B) \times (A' \times B)\).
+    We choose its restriction.
+
+    Is this restriction provably functional? The only potential point of failure is the existence of an output value. So we must show:
+    \[(\vec{a},\vec{b}) : (A*B) \vdash \exists (\vec{a'},\vec{b'}) : (A'*B), (f*1_B)((\vec{a},\vec{b}), (\vec{a'},\vec{b'}))\]
+
+    Or equivalently:
+    \[\vec{a} : A, \vec{b} : B, \vec{a} \sim \vec{b} \vdash \exists \vec{a'} : A', \exists \vec{b'} : B, \vec{a'} \sim \vec{b'} \land f(\vec{a},\vec{a'}) \land \vec{b} = \vec{b'}\]
+    \[\vec{a} : A, \vec{b} : B, \vec{a} \sim \vec{b} \vdash \exists \vec{a'} : A', \vec{a'} \sim \vec{b} \land f(\vec{a},\vec{a'})\]
+
+    But this is obtained from \(\vec{a} : A \vdash \exists \vec{a'} : A', f(\vec{a},\vec{a'})\) by applying Lemma \ref{transform-lemma} once for each variable in \(B\).
+    And that holds because \(f\) is provably functional.
+
+    \vspace{1em}
+
+    We've defined \(f * 1_B\). The definition of \(1_A * g\) is analogous.
+    All of the functoriality laws, including the interchange condition \((f * 1) \circ (1 * g) = (1 * g) \circ (f * 1)\),
+    follow from relating \(f * 1_B\) back to \(f \times 1_B\), and using the bifunctoriality of the cartesian product.
+\end{proof}
+
+\begin{lemma}\leavevmode
+    \begin{enumerate}
+        \item The monoidal structure is symmetric, with the terminal object as the monoidal unit.
+        \item The canonical map \(A*B \to A \times B\) is a monomorphism.
+        \item For any \(X\), the endofunctor \(-*X\) preserves pullbacks and covers.
+    \end{enumerate}
 \end{lemma}
 
+\begin{proof}\leavevmode
+    \begin{enumerate}
+        \item Immediate by chasing definitions.
+        \item \(A*B\) is simply \(A \times B\) with extra hypotheses.
+        \item \begin{description}
+            \item[Pullbacks:] Consider any cospan \(A \xrightarrow{f} C \xleftarrow{g} B\). We have:
+            \begin{align*}
+                A \times_C B &= \vec{a} : A, \vec{b} : B, \vec{c} : C, f(a,c), g(b,c)
+                \\ (A \times_C B) * X &= \vec{a} : A, \vec{b} : B, \vec{c} : C, f(a,c), g(b,c),
+                \\ &\qquad \vec{x} : X, \vec{a} \sim \vec{x}, \vec{b} \sim \vec{x}, \vec{c} \sim \vec{x}
+                \\ (A*X) \times_{C*X} (B*X) &= \vec{a} : (A*X), \vec{b} : (B*X), \vec{c} : (C*X),
+                \\ &\qquad (f*1_X)(a,c), (g*1_X)(b,c)
+                \\ &= \vec{a} : A, \vec{x}_1 : X, \vec{a} \sim \vec{x}_1,
+                \\ &\qquad \vec{b} : A, \vec{x}_2 : X, \vec{b} \sim \vec{x}_2,
+                \\ &\qquad \vec{c} : A, \vec{x}_3 : X, \vec{c} \sim \vec{x}_3,
+                \\ &\qquad f(a,c), \vec{x}_1 = \vec{x}_3,
+                \\ &\qquad g(b,c), \vec{x}_2 = \vec{x}_3
+            \end{align*}
+            \item[Covers:] Consider any cover \(\{f_i : A_i \to B | i \in I\}\).
+            Since this is a cover:
+            \[\vec{b} : B \vdash \bigvee_{i \in I} \exists \vec{a} : A_i, f_i(\vec{a},\vec{b})\]
+            We are to show that \(\{f_i*1_X : A_i*X \to B*X | i \in I\}\) is a cover. That is:
+            \[\vec{b} : B*X \vdash \bigvee_{i \in I} \exists \vec{a} : A_i*X, (f_i*1_X)(\vec{a},\vec{b})\]
+            Equivalently:
+            \[\vec{b} : B, \vec{x} : X, \vec{b} \sim \vec{x} \vdash \bigvee_{i \in I} \exists \vec{a} : A_i, \exists \vec{y} : X, \vec{a} \sim \vec{y} \land f_i(\vec{a},\vec{b}) \land \vec{y} = \vec{x}\]
+            \[\vec{b} : B, \vec{x} : X, \vec{b} \sim \vec{x} \vdash \bigvee_{i \in I} \exists \vec{a} : A_i, \vec{a} \sim \vec{x} \land f_i(\vec{a},\vec{b})\]
+            But this follows from applying the transformation in Lemma \ref{transform-lemma} to the hypothesis, once for each variable in \(X\).
+        \end{description}
+    \end{enumerate}
+\end{proof}
+
+Lift to the classifying topos.
+
+\begin{construction} \label{sh}
+    We equip \(\Set[\T]\) with a strictly affine symmetric monoidal closed structure \((*,\wand)\).
+\end{construction}
+
 \begin{proof}
-    \(\Sh(C,J)\) is a Grothendieck topos, hence a topos. There is a canonical choice of pullbacks in \(\Sh(C,J)\), given by pointwise lifting the canonical pullbacks in \(C\).
+    Let \(\C_\T\) be \(\T's\) syntactic category. Recall that \(\Set[\T] = \Sh(\C_\T)\).
 
-    Now, theorem 4.3.2 from \cite{biering} shows the existence of a monidal structure \(\otimes^\Sh\) on \(\Sh(C,J)\),
-    given by transporting Day convolution across the adjunction \((\mathbf{a} \dashv i) : \Psh(C,J) \rightleftarrows \Sh(C,J)\).
+    Theorem 4.3.2 from \cite{biering} defines a monidal structure \(\otimes^\Sh\) on \(\Sh(\C_\T)\),
+    given by transporting Day convolution across the adjunction \((\mathbf{a} \dashv i) : \Psh(\C_\T) \rightleftarrows \Sh(\C_\T)\).
     Corollary 4.3.10 from the same shows that it is closed.
-    Chasing definitions quickly shows \(\otimes^\Sh\) is symmetric.
-    And the monoidal unit is the sheafification of the Yoneda embedding of \(*\)'s unit;
-    since \(*\) is affine, this simplifies to the terminal object.
 
-    In other words, \(\otimes^\Sh\) is an affine symmetric monoidal closed structure.
-    It remains to show that this is strict --- that the canonical map \(A \otimes^\Sh B \to A \times B\) in \(\Sh(C,J)\) is always a monomorphism.
+    The monoidal unit is the sheafification of the Yoneda embedding of \(*\)'s unit.
+    Since \(*\) is affine, this simplifies to the terminal object.
 
-    We can reduce to the case of presheaves.
+    Symmetry is a simple definition chase. So it remains only to show that
+    that the canonical map \(A \otimes^\Sh B \to A \times B\) in \(\Sh(C,J)\) is always a monomorphism.
+
     Letting \(\otimes^\Psh\) represent Day convolution on \(\Psh(C)\),
-    the map of interest is just the result of applying the sheafification functor, which preserves monomorphisms, to the canonical map \(iA \otimes^\Psh iB \to iA \times iB\).
+    the map of interest is just the result of applying the sheafification functor,
+    which preserves monomorphisms, to the canonical map \(iA \otimes^\Psh iB \to iA \times iB\).
 
     Expanding both \(\otimes^\Psh\) and \(\times\) as Day convolutions,
-    this is a set-map, natural in \(c \in C^\mathrm{op}\), of the following type.
+    this is a \(\Set\)-map, natural in \(c \in C^\mathrm{op}\), of the following type.
     \begin{scriptsize}\[\left(\int^{c_1,c_2 \in C} A(c_1) \times B(c_2) \times (c \xrightarrow{C} c_1 * c_2)\right) \to \left(\int^{c_1,c_2 \in C} A(c_1) \times B(c_2) \times (c \xrightarrow{C} c_1 \times c_2)\right)\]\end{scriptsize}
@@
 
-Finally, we set the site to a syntactic site, to understand classifying toposes.
+And finally:
 
-\begin{lemma}
-    Let \(\T\) be a geometric theory, satisfying the conditions of the main theorem.
-    Then the syntactic site for \(\T\) satisfies the conditions of the above lemma.
-\end{lemma}
+\begin{construction}
+    We have an interpretation of \(\Set[\T]\) into \(\mathbf{BT}(*, 1, \Sigma, \Pi, \Pi^*)\).
+\end{construction}
 
 \begin{proof}
-    Let us define a few shorthands.
-
-    We'll allow ourselves to write sequents involving compound types, such as \(p : A \times B \vdash \pi_1 p \sim \pi_2 p\).
-    These can be straightforwardly ``compiled out'' to ordinary sequents, such as \(a : A, b : B \vdash a \sim b\).
-
-    Next, if we have a context \(\Gamma = (a_1 : A_1, \dots, a_n : A_n, \phi_1 \dots \phi_p)\),
-    we'll reuse the name \(\Gamma\) for the type \(\{(a_1, \dots, a_n) : A_1 \times \dots \times A_n \mid \phi_1 \land \dots \land \phi_p\}\).
-    If we have a substitution \(f : \Gamma \to \Delta\) and a term \(\gamma : \Gamma\),
-    there is a natural way to define the term \(f\gamma : \Delta\).
-
-    Finally, if \(\gamma = (a_1, \dots, a_m) : \Gamma\) and \(\delta = (b_1, \dots, b_n) : \Delta\),
-    we write \(\gamma \sim \delta\) as a shorthand for \(\bigwedge_{i=1}^m \bigwedge_{j=1}^n a_i \sim b_j\).
-
-    With that out of the way, let's begin the proof.
-
-    \vspace{1em}
-
-    \noindent The syntactic category is guaranteed to be finitely complete, with a canonical construction for pullbacks.
-    Specifically, the empty context is terminal,
-    and the pullback of the span \(\Gamma \xrightarrow{f} \Xi \xleftarrow{g} \Delta\) is the context \((\gamma : \Gamma, \delta : \Delta, f\gamma = g\delta)\).
-
-    Given contexts \(\Gamma\) and \(\Delta\), define \(\Gamma * \Delta = (\gamma : \Gamma, \delta : \Delta, \gamma \sim \delta)\).
-    I claim this is a strict affine symmetric monoidal product on the syntactic category.
-    \begin{itemize}
-        \item Associativity and the unit laws are easily checked, and symmetry reduces to the symmetry of \(\sim\).
-        \item Functoriality says that if \(f : \Gamma \to \Gamma'\), \(f \times \mathrm{id}_\Delta\) is a valid map \(\Gamma * \Delta \to \Gamma' * \Delta\).
-        This reduces to the claim \(x : \Delta, \gamma : \Gamma, \gamma \sim x \vdash f\gamma \sim x\).
-        Further reducing to the same claim about the basic sorts and function symbols, this is exactly one of the hypotheses, as detailed in Section \ref{Details}.
-        \item Since \(\Gamma * \Delta\) is a subtype of \(\Gamma \times \Delta\), strictness holds, and the coherence equations are trivial.
-    \end{itemize}
-
-    It remains to show that for any context \(\Theta\), the functor \(- * \Theta\) preserves pullbacks and covers.
-
-    Preservation of pullbacks is a direct computation. For the span \(\Gamma \xrightarrow{f} \Xi \xleftarrow{g} \Delta\),
-    it simply says the following two contexts are isomorphic, a fact which is easily checked.
-    \begin{align*}
-        &(\gamma : \Gamma, \delta : \Delta, f\gamma = g\delta, \theta : \Theta, \gamma \sim \theta, \delta \sim \theta) \\
-        &(\gamma : \Gamma, \theta_1 : \Theta, \gamma \sim \theta_1, \delta : \Delta, \theta_2 : \Theta, \delta \sim \theta_2, f\gamma = g\delta, \theta_1 = \theta_2)
-    \end{align*}
-
-    That leaves covers. Suppose \(\{f_i : \Gamma_i \to \Delta \mid i \in I\}\) is a cover,
-    meaning \(\delta : \Delta \vdash \bigvee_{i \in I} \exists \gamma : \Gamma_i, f_i\gamma = \delta\) is \(\T\)-provable.
-
-    We know by assumption that for any sort \(X\) and any \(x : X\),
-    \(\T\)-provable statements remain \(\T\)-provable upon restricting each sort \(A\) to its subset \(\{a : A \mid a \sim x\}\).
-    Iterating this fact, the same is true if we replace the sort \(X\) by a context, such as \(\Theta\).
-
-    Therefore, we know the following statement is provable.
-    \[\theta : \Theta, \delta : \{\delta : \Delta \mid \delta \sim \theta\} \vdash \bigvee_{i \in I} \exists \gamma : \{\gamma : \Gamma_i \mid \gamma \sim \theta\}, f_i\gamma = \delta\]
-
-    But that's equivalent to the following.
-    \[\delta : (\Delta * \Theta) \vdash \bigvee_{i \in I} \exists \gamma : (\Gamma_i * \Theta), (f_i * \mathrm{id}_\Theta)\gamma = \delta\]
-
-    So \(\{f_i * \mathrm{id}_\Theta : \Gamma_i * \Theta \to \Delta * \Theta \mid i \in I\}\) is a cover, as required.
-\end{proof}
-
-Putting everything together:
-
-\begin{proof}[Proof of Main Theorem]
-    Combine the above four lemmas.
+    Apply section 3.5 of \cite{schopp} to see that \(\Set[\T]\) is a ``\((*, 1, \Sigma, \Pi, \Pi^*)\)-type-category'',
+    then apply Chapter 6 of the same to construct the interpretation.
 \end{proof}
@@
 \begin{theorem}
-    Recall that \(\Set[\T]\) contains the generic model of \(\T\).
-    Suppose \(A\) and \(B\) are sorts of \(\T\), interpreted in the generic model.
-    Then, internally to the language, a pair \((a,b)\) is in the image of the inclusion \(A * B \hookrightarrow A \times B\) exactly when \(a \sim b\).
+    Let \(U\) be the generic \(\T\)-model in \(\Set[\T]\).
+    Then \(U * U = \{(a,b) : U \times U \mid a \sim b\}\), as subsets of \(U \times U\).
 \end{theorem}
@@
 \begin{proof}
-    To avoid confusion, write \(A\) to refer to the sort in \(\T\),
-    \((a:A)\) to refer to the \emph{context} in \(\T\),
-    and \(U_A\) to refer to the corresponding type in the generic model.
+    Viewing \(\Set[\T]\) as \(\Sh(\C_\T)\), expand.
+    \begin{align*}
+        U \otimes^\Sh U
+        &\simeq \mathbf{a}\yo(x:X) \otimes^\Sh \mathbf{a}\yo(x:X)
+        \\ &\simeq \mathbf{a}(i\mathbf{a}\yo(x:X) \otimes^\Psh i\mathbf{a}\yo(x:X))
+        \\ &\simeq \mathbf{a}(\yo(x:X) \otimes^\Psh \yo(x:X))
+        \\ &\simeq \mathbf{a}\yo((x:X) * (x:X))
+        \\ &\simeq \mathbf{a}\yo(a:X, b:X, a \sim b)
+        \\ &\simeq \{(a,b) : U \times U \mid a \sim b\}
+    \end{align*}
 
-    Then \(U_A\) interprets as \(\mathbf{a}\yo(a:A)\), and similarly for \(B\).
-    The type \(\{(a,b) : U_A \times U_B \mid a \sim b\}\) interprets as \(\mathbf{a}\yo(a:A, b:B, a \sim b)\),
-    and its inclusion into \(U_A \times U_B\) is \(\mathbf{a} \circ \yo\) applied to the obvious substitution \((a:A,b:B,a \sim b) \to (a:A,b:B)\).
-
-    Now, in the proof of Lemma \ref{sh}, we computed the interpretation of a general inclusion \(X * Y \hookrightarrow X \times Y\).
-    Specializing to the types \(U_A\) and \(U_B\), that result simplifies to the obvious map
-    \(\mathbf{a}\yo((a:A) * (b:B)) \to \mathbf{a}\yo((a:A) \times (b:B))\).
-    That is, it's \(\mathbf{a} \circ \yo\) applied to the obvious substitution from \((a:A,b:B,a \sim b)\) to \((a:A,b:B)\).
-
-    So the interpretations match. Thus \(U_A * U_B = \{(a,b) : U_A \times U_B \mid a \sim b\}\), as subsets of \(U_A \times U_B\).
+    Tracing the first projection \(1_U \otimes^\Sh ! : U \otimes^\Sh U \to U \otimes^\Sh \mathbf{1}\)
+    through the chain of isomorphisms, we see that it becomes the map \((a,b) \mapsto a\)
+    from \(\{(a,b) : U \times U \mid a \sim b\}\) to \(U\).
+    Similarly for the second projection.
+    The canonical map \(U \otimes^\Sh U \to U \times U\) is the product of these,
+    so it becomes the inclusion \(\{(a,b) : U \times U \mid a \sim b\} \hookrightarrow U \times U\),
+    as required.
 \end{proof}

### 2025-06-29 paper.tex
Bare bones argument for reverse direction. Needs formatting and reorganization.

@@
 
-\usepackage[style=alphabetic]{biblatex}
-\addbibresource{paper.bib}
+% \usepackage[style=alphabetic]{biblatex}
+% \addbibresource{paper.bib}
 
@@
 \newtheorem{lemma}[theorem]{Lemma}
+\newtheorem{corollary}[theorem]{Corollary}
 \newtheorem{construction}[theorem]{Construction}
@@
 
+\section{Backwards}
+
+TODO: Trying to make the main theorem an iff statement
+TODO: This entire section will need to be rewritten once the proof works.
+POSSIBILITY: Make the proof a chain of iffs. That way, there's a non-classifying characterization of topos models as well as a classifying one.
+% culminating in a beautiful characterization in terms of classifying toposes.
+
+Consider any model of \(\mathbf{BT}(*, 1, \Sigma, \Pi, \Pi^*)\) in the classifying topos over \(\T\).
+
+TODO: Translate properties of the type theory to properties of the topos. E.g. \(*\) affine monoidal product...
+
+Let \(U\) be the generic \(\T\)-model in \(\Set[\T]\).
+
+\begin{lemma}
+    For any \(A \in \Set[\T]\), the data \(A * -, \Pi^*_A\) is determined by the subset \(A * U \subseteq A \times U\).
+\end{lemma}
+
+\begin{proof}
+    Observe that TODO form a geometric morphism \(\Set[\T]_{/A} \to \Set[\T]\).
+    Each such geometric morphism yields a \(\T\)-model in \(\Set[\T]_{/A}\),
+    given by the restriction of the obvious model \(\substack{A \times U \\ \downarrow \\ A}\) to \(\substack{A * U \\ \downarrow \\ A}\).
+\end{proof}
+
+\begin{corollary}
+    The monoidal structure is determined by \(U * U\).
+\end{corollary}
+
+\begin{proof}
+    Reduce from \(A*B\) to \(A*U\), then to \(U*U\).
+\end{proof}
+
+Next, we relate the classifying topos back to the syntactic category.
+
+\begin{lemma}
+    If an object of \(\Set[\T]\) is the interpretation of some context \(A \in \C_\T\) in the generic model,
+    then every subobject is the interpretation of some restriction \((\vec{a} : A, \phi(\vec{a}))\).
+    The choice of \(\phi\) is unique up to logical equivalence.
+\end{lemma}
+
+\begin{proof}
+    Writing \(\Omega\) for the subobject classifier of \(\Set[\T]\), we have:
+    \begin{align*}
+        &\mathrm{Sub}(\mathbf{a}\yo A)
+        \\ &= \mathbf{a}\yo A \to \Omega
+        \\ &= \hom_{\Sh(\C_\T)}(\mathbf{a}\yo A, \Omega)
+        \\ &= \hom_{\Psh(\C_\T)}(\yo A, i\Omega)
+        \\ &= \Omega A
+    \end{align*}
+    This is the collection of covering-closed sieves on \(A\) in the syntactic category.
+
+    A sieve has the required form when it contains exactly those morphisms that factor through the inclusion of \((\vec{a}:A,\phi(a))\) into \(A\).
+    Uniqueness is immediate; if \(\phi_1\) and \(\phi_2\) are both solutions,
+    then the inclusions of \((\vec{a}:A,\phi_1(a,b))\) and \((\vec{a}:A,\phi_2(a,b))\) factor through each other, producing a bi-implication.
+
+    As for existince, fix any sieve \(S\).
+    Every morphism it contains factors through the inclusion for the following choice of \(\phi\).
+    \[\phi(\vec{a}) = \bigvee_{(B \in \C_\T, f : \hom_{\C_\T}(B, A)) \in S} \exists \vec{b} : B, f(\vec{b},\vec{a})\]
+    And conversely, this inclusion is covered by the sieve, so covering-closedness ensures it it in the sieve.
+\end{proof}
+
+\begin{corollary}
+    A geometric sequent \(\vec{a} : A \vdash \phi(\vec{a})\) is provable if it holds in the generic model.
+\end{corollary}
+
+\begin{proof}
+    Uniqueness, from the above lemma, shows that \(\phi\) and \(\top\) are logically equivalent.
+\end{proof}
+
+\begin{corollary}
+    A morphism in \(\Set[\T]\) is the interpretation of one in the generic model if its domain and codomain are.
+\end{corollary}
+
+\begin{proof}
+    A morphism in \(\Set[\T]\) induces a provably functional relation.
+    Both the relation and the proof of functionality are interpretations of the same in the syntactic category.
+    This yields the data required to produce a morphism in \(\C_\T\).
+\end{proof}
+
+\begin{theorem}
+    \(U * U\) is the interpretation of a unique geometrically-definable relation \((\sim) : X \times X \to \Prop\).
+    This relation satisfies the assumptions of the main theorem.
+\end{theorem}
+
+\begin{proof}
+    \(U * U\) is a subobject of \(U \times U\), which interprets \(X \times X\).
+    So \(U \times U\) is the interpretation of some context \((a:X,b:X,a\sim b)\).
+
+    It remains to show that \((\sim)\) is symmetric,
+    and that \(\{a : X \mid a \sim x\}\) is a sub-\(\T\)-model of \(X\) for any \(x : X\).
+
+    By the above lemma and its corollaries, it suffices to show that \((a,b) \in U*U \Leftrightarrow (b,a) \in (U*U)\),
+    and that \(\{a : U \mid a \sim x\}\) is a sub-\(\T\)-model of \(U\) for any \(x : U\).
+
+    The former is immediate. The latter holds because the geometric morphism TODO induces a \(\T\)-model in \(\Set[\T]_{/A}\).
+\end{proof}
+
 \section{Future Work}
@@
 
-\printbibliography
+% \printbibliography
 

### 2025-06-30 paper.tex
Clean up proof. Completes handling of issue #2.

@@
 
-% \usepackage[style=alphabetic]{biblatex}
-% \addbibresource{paper.bib}
+\usepackage[style=alphabetic]{biblatex}
+\addbibresource{paper.bib}
 
@@
 \usepackage{mathpartir}
+\usepackage{quiver}
+
+\usepackage{todonotes}
 
@@
     In their PhD thesis, Sch\"opp introduces a type theory designed to capture the monoidal structure in the Schanuel topos.
-    I provide a simple condition that can be used to verify whether the type theory interprets into a classifying topos of your choice.
+    I provide a strikingly simple characterization of the models of this type theory in arbitrary classifying toposes.
 \end{abstract}
+
+\listoftodos
 
@@
 
-I provide an easy method to verify that \(\mathbf{BT}(*, 1, \Sigma, \Pi, \Pi^*)\)
-can be interpreted in a classifying topos of your choice,
-so you can start \emph{using} the internal language as quickly as possible.
+I characterize the models of \(\mathbf{BT}(*, 1, \Sigma, \Pi, \Pi^*)\) in an arbitrary classifying topos.
+This characterization is extremely easy to use; describing a topos model typically becomes a single-paragraph task.
+
+For simplicity, I describe only the case of single-sorted type theories.
+While both the theorem and proof extend naturally to multi-sort theories,
+the notation becomes more awkward in that case.
 
@@
 
-For simplicity of exposition, I describe the theorem and proof in the case of single-sorted theories.
-The extension to multiple sorts is straightforward, but terminologically awkward.
+Recall from \cite{schopp}, section 3.5 and chapter 6, that topos models of the theories
+\(\mathbf{BT}(1, \Sigma, \Pi, \Pi^*)\), \(\mathbf{BT}(*, 1, \Sigma, \Pi, \Pi^*)\), or \(\mathbf{BT}(*, 1, \Sigma, \Pi, \Pi^*, B^{*(M:A)})\)
+come from toposes equipped with \emph{strict affine symmetric monoidal closed structures that preserve pullback}.
+That is, symmetric monoidal structures \((*,I)\) where:
+\begin{itemize}
+    \item \(I\) is terminal.
+    \item The canonical map \(A * B \to A \times B\) is mono.
+    \item The functor \(A * -\) preserves pullbacks, and has a right adjoint.
+\end{itemize}
+
+I characterize these structures in classifying toposes.
 
@@
     Let \(\T\) be a single-sorted geometric theory, with sort \(X\).
-    Let \((\sim) : X \to X \to \Prop\) be a geometrically-defined symmetric relation.
-    Further suppose that for any \(x \in X\), \(\{a : X \mid a \sim x\}\) is a sub-\(\T\)-model of \(X\).
-    Then \(\mathbf{BT}(*, 1, \Sigma, \Pi, \Pi^*)\) interprets into \(\Set[\T]\).
+    Monoidal structures on topos \(\Set[\T]\) of the above form 
+    are given by symmetric relations \((\sim) : X \to X \to \Prop\),
+    satisfying the following property.
+    \begin{quote}
+        For any \(x : X\), the \(\T\)-model structure on \(X\)
+        restricts to a sub-\(\T\)-model on \(\{a : X \mid a \sim x\}\).
+    \end{quote}
+
+    Under this interpretation, if \(U\) is the generic \(\T\)-model,
+    we will have \(U*U = \{(a,b) : U \times U \mid a \sim b\}\).
 \end{theorem}
 
-For example, let \(\T\) be the theory of rings, and let \(a \sim b\) mean \(ab = ba\).
-The hypothesis holds because centralizers of rings are rings;
-so we get an interpretation of the bunched type theory into \(\Set[\Ring]\).
+\noindent For example, to interpret the type theory into \(\Set[\Ring]\), let \(a \sim b\) mean \(ab = ba\).
+We need only check that \(\{a : R \mid ax = xa\}\) is a ring for any \(x\), and this holds by standard ring theory.
+The monoidal product of the generic ring with itself will be the set of commuting pairs of ring elements.
+
+% \begin{theorem}
+%     Let \((\sim) : X \to X \to \Prop\) be a symmetric relation, geometrically defined in the language of the theory.
+%     Further suppose that for any \(x \in X\), \(\{a : X \mid a \sim x\}\) is a sub-\(\T\)-model of \(X\).
+%     Then \(\mathbf{BT}(*, 1, \Sigma, \Pi, \Pi^*)\) interprets into \(\Set[\T]\).
+% \end{theorem}
+
+% For example, if \(\T\) is the theory of rings, let \(a \sim b\) mean \(ab = ba\).
+% The hypothesis holds because centralizers of rings are rings,
+% so we get an interpretation of the bunched type theory into \(\Set[\Ring]\).
 
@@
 
-As an example of the latter, say \(\T\) is the theory of inhabited sets, with axiom \(\vdash \exists a : X, \top\).
+As an example of the latter condition, say \(\T\) is the theory of inhabited sets, with axiom \(\vdash \exists a : X, \top\).
 Then the required sequent is \(x : X \vdash \exists a : X, a \sim x \land \top\).
@@
 
-For the duration of this section, we accept the assumptions of the main theorem.
+\subsection{Notation}
 
-\vspace{1em}
+We use the following notation.
+\begin{itemize}
+    \item \(\T\) will be a single-sorted geometric theory, with sort \(X\).
+    \item \(\C_\T\) will be the syntactic site of \(\T\).
+    \item \(\Psh(\C_\T)\) and \(\Sh(\C_\T)\) are the categories of presheaves and sheaves on the syntactic site.
+    The latter is the classifying topos of \(\T\), so we also call it \(\Set[\T]\).
+    \item We have the following diagram of functors:
+    % https://q.uiver.app/#q=WzAsMyxbMCwwLCJcXENfXFxUIl0sWzIsMCwiXFxQc2goXFxDX1xcVCkiXSxbNCwwLCJcXFNoKFxcQ19cXFQpID0gXFxTZXRbXFxUXSJdLFswLDEsIlxceW8iXSxbMSwyLCJcXG1hdGhiZnthfSIsMCx7Im9mZnNldCI6LTJ9XSxbMiwxLCJpIiwwLHsib2Zmc2V0IjotMn1dLFs0LDUsIiIsMCx7ImxldmVsIjoxLCJzdHlsZSI6eyJuYW1lIjoiYWRqdW5jdGlvbiJ9fV1d
+    \[\begin{tikzcd}
+        {\C_\T} && {\Psh(\C_\T)} && {\Sh(\C_\T) = \Set[\T]}
+        \arrow["\yo", from=1-1, to=1-3]
+        \arrow[""{name=0, anchor=center, inner sep=0}, "{\mathbf{a}}", shift left=2, from=1-3, to=1-5]
+        \arrow[""{name=1, anchor=center, inner sep=0}, "i", shift left=2, from=1-5, to=1-3]
+        \arrow["\dashv"{anchor=center, rotate=-90}, draw=none, from=0, to=1]
+    \end{tikzcd}\]
+    \(\yo\) is the Yoneda embedding, \(\mathbf{a}\) is sheafification, and \(i\) is the inclusion of presheaves into sheaves.
+    The composite functor \(\mathbf{a} \circ \yo\) interprets the syntax of the theory into the generic model.
+    \item \(U = \mathbf{a}\yo(x:X)\) is the interpretation of the \(\T\)'s sort in the generic model.
+\end{itemize}
 
-\noindent To begin, notice a key detail of the transformation described above.
+\subsection{Forwards Direction}
+
+For the duration of this section, let \((\sim) : X \to X \to \Prop\) be a relation satisfying the property described in the main theorem.
+
 \begin{lemma} \label{transform-lemma}
@@
 
-Next, bootstrap to the syntactic site.
 \begin{construction}
@@
     All of the functoriality laws, including the interchange condition \((f * 1) \circ (1 * g) = (1 * g) \circ (f * 1)\),
-    follow from relating \(f * 1_B\) back to \(f \times 1_B\), and using the bifunctoriality of the cartesian product.
+    follow from relating \(f * 1_B\) back to \(f \times 1_B\), and using the functoriality of the cartesian product.
 \end{proof}
@@
 
-Lift to the classifying topos.
-
 \begin{construction} \label{sh}
-    We equip \(\Set[\T]\) with a strictly affine symmetric monoidal closed structure \((*,\wand)\).
+    We equip \(\Set[\T]\) with a pullback-preserving strict affine symmetric monoidal closed structure.
 \end{construction}
@@
     Since that canonical map is mono, this is too.
+
+    \todo{Missing condition: Pullback preservation. I'm not sure how to prove this, or if I need to change the theorem statement to make it work.}
 \end{proof}
-
-And finally:
-
-\begin{construction}
-    We have an interpretation of \(\Set[\T]\) into \(\mathbf{BT}(*, 1, \Sigma, \Pi, \Pi^*)\).
-\end{construction}
-
-\begin{proof}
-    Apply section 3.5 of \cite{schopp} to see that \(\Set[\T]\) is a ``\((*, 1, \Sigma, \Pi, \Pi^*)\)-type-category'',
-    then apply Chapter 6 of the same to construct the interpretation.
-\end{proof}
-
-\section{Using the Monoidal Structure}
-
-For practical use, the main theorem isn't quite enough.
-You'll need to know not just that \(*\)-types exist, but how they work.
-If \(R\) is the generic local ring, you'll want to know that \(R * R = \{(a,b) : R^2 \mid ab = ba\}\).
-
-Hence this next theorem.
 
@@
     Let \(U\) be the generic \(\T\)-model in \(\Set[\T]\).
-    Then \(U * U = \{(a,b) : U \times U \mid a \sim b\}\), as subsets of \(U \times U\).
+    Then \(U \otimes^\Sh U = \{(a,b) : U \times U \mid a \sim b\}\), as subsets of \(U \times U\).
 \end{theorem}
@@
 \begin{proof}
-    Viewing \(\Set[\T]\) as \(\Sh(\C_\T)\), expand.
     \begin{align*}
@@
 
-\section{Backwards}
+\subsection{Backwards Direction}
 
-TODO: Trying to make the main theorem an iff statement
-TODO: This entire section will need to be rewritten once the proof works.
-POSSIBILITY: Make the proof a chain of iffs. That way, there's a non-classifying characterization of topos models as well as a classifying one.
-% culminating in a beautiful characterization in terms of classifying toposes.
-
-Consider any model of \(\mathbf{BT}(*, 1, \Sigma, \Pi, \Pi^*)\) in the classifying topos over \(\T\).
-
-TODO: Translate properties of the type theory to properties of the topos. E.g. \(*\) affine monoidal product...
-
-Let \(U\) be the generic \(\T\)-model in \(\Set[\T]\).
+Consider any pullback-preserving strict affine symmetric monoidal structure on \(\Set[\T]\).
 
 \begin{lemma}
-    For any \(A \in \Set[\T]\), the data \(A * -, \Pi^*_A\) is determined by the subset \(A * U \subseteq A \times U\).
+    For any \(A \in \Set[\T]\), there is a geometric morphism \(W_A \dashv \Pi^*_A : \Set[\T]_{/A} \leftrightarrows \Set[\T]\),
+    whose composition with \(\Sigma_A \dashv A^* : \Set[\T] \leftrightarrows \Set[\T]_{/A}\)
+    yields the adjunction \((A * -) \dashv (A \wand -)\).
 \end{lemma}
@@
 \begin{proof}
-    Observe that TODO form a geometric morphism \(\Set[\T]_{/A} \to \Set[\T]\).
-    Each such geometric morphism yields a \(\T\)-model in \(\Set[\T]_{/A}\),
+    Following the construction in section 2.5.1 of \cite{schopp}, we define:
+    \begin{align*}
+        W_A(B) &= \substack{A * B \\ \downarrow \\ A}
+        \\ \Pi^*_A\left(\substack{B \\ \enspace \downarrow f \\ A}\right) &= \text{Pullback of } (A \wand B \xrightarrow{A \wand f} A \wand A \xleftarrow{id} 1)
+    \end{align*}
+\end{proof}
+
+\begin{corollary}
+    For any \(A \in \Set[\T]\), the functor \(A * -\) is determined by \(A * U\).
+\end{corollary}
+
+\begin{proof}
+    The above geometric morphism corresponds to a \(\T\)-model in \(\Set[\T]_{/A}\),
     given by the restriction of the obvious model \(\substack{A \times U \\ \downarrow \\ A}\) to \(\substack{A * U \\ \downarrow \\ A}\).
@@
 
-    The former is immediate. The latter holds because the geometric morphism TODO induces a \(\T\)-model in \(\Set[\T]_{/A}\).
+    The former is immediate. The latter holds because the geometric morphism \(W_U \dashv \Pi^*_U\) induces a \(\T\)-model in \(\Set[\T]_{/U}\).
 \end{proof}
@@
 
-% \printbibliography
+\printbibliography
 
